---
title: "ACT"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(extrafont)
library(readxl)
library(janitor)
library(lubridate)

loadfonts()

```

<br>

# Prep Data

Next up is to join the ACT data. 

<br>

```{r act original data}
act.original <- read_csv("Data/SLEDS/ACT/ACT.csv") %>%
  mutate(PersonID = as.integer(PersonID))

head(act.original)

names(act.original)
```

<br>

There are `r comma(nrow(act.original), accuracy = 1)` observations in the dataset. Each observation is an ACT test taken by a student in Southwest Minnesota. The unique feature here is the time it was taken. This means that a student could have taken the ACT twice and their PersonID would show up as two observations.

Out of the scores provided, I'm just going to use the composite score. Since an individual may have taken the test multiple times I will only use the highest composite score for each individual. This is how colleges do it as well.

<br>

```{r act highest composite score}
act.highest.comp <- act.original %>%
  select(PersonID, ACTCompositeScore) %>%
  mutate(PersonID = as.integer(PersonID)) %>%
  group_by(PersonID) %>%
  filter(ACTCompositeScore == max(ACTCompositeScore)) %>%
  ungroup() %>%
  distinct(PersonID, .keep_all = TRUE)

head(act.highest.comp)

names(act.highest.comp)
```

<br>

After filtering for only the highest ACT composite score for each PersonID, the dataset contains `r comma(nrow(act.highest.comp), accuracy = 1)` observations.

Now we will join it up with the master.

<br>

```{r act join with master}
master.2 <- read_csv("Data/SLEDS/Masters/master-1.csv") %>%
  left_join(act.highest.comp, by = "PersonID")

head(master.2)

names(master.2)

```

<br>

The updated master dataset contains `r comma(nrow(master.2), accuracy = 1)` observations with `r comma(ncol(master.2), accuracy = 1)` columns.

<br>

# Summary of ACT data

