---
title: "Analysis - multiple correspondance analysis"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(readxl)
library(janitor)
library(lubridate)
library(systemfonts)
reset_font_cache()
library(ggtext)
library(gmodels)
library(leaps)
library(bestglm)
library(fastDummies)
library(car)
library(FactoMineR)
library(factoextra)
library(psych)
library(corrplot)
```

```{r join docs, include=FALSE}
theme_bar <- theme_bw() +
  theme(panel.grid.major = element_line(color = "grey70", size = 0.1),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.key.size = unit(1, "lines"),
        legend.margin = margin(0,0,0,0),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        text = element_text(family = "Arial") ,
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"))

theme_line <- theme_bw() +
  theme(legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.text = element_text(margin = margin(l = 2)),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey70", size = 0.1),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial") ,
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"))


theme_sf <- theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "white"),
        panel.border = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial") ,
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"))

regions <- read_csv("Data/Join docs/county_regions.csv") %>%
    select(5,6) %>%
    unique() %>%
    mutate(edr = str_replace(edr, "  ", " "),
           planning.region = str_replace(planning.region, " Minnesota", ""),
           planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
           edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"))

counties.regions <- read_csv("Data/Join docs/county_regions.csv") %>%
  rename(mif = `MIF Region`) %>%
  mutate(countyfp = formatC(countyfp, width = 3, flag = "0"),
         Name = str_to_title(Name),
         Name = str_replace(Name, "Q", "q"),
         Name = str_replace(Name, "Of The", "of the"),
         Name = str_replace(Name, "Mcleod", "McLeod"),
         Dem_Desc = ifelse(Name == "Minnesota", "Minnesota", Dem_Desc) ,
         edr = str_replace(edr, "  ", " "),
         planning.region = str_replace(planning.region, " Minnesota", ""),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"),
         mif = ifelse(is.na(mif), "TC", mif),
         mif = as.factor(mif),
         mif = fct_relevel(mif, "NW", "NE", "WC", "EC", "SW", "SE", "TC"))


color.ruca <- c("Entirely rural" = "#009933", "Town/rural mix" = "#99CC33", "Urban/town/rural mix" = "#CC9966", "Entirely urban" = "#754C29", "Minnesota" = "black")

color.pr <- c("Northwest" = 	"#4575b4", "Northeast" = "grey", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black")

color.edr <- c("EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365", "Minnesota" = "black")

color.pr.edr <- c ("Northwest" = "#4575b4","Northeast" = "#e0f3f8", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black", "EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365")

mn_counties <- st_read("Data/Shapefiles/county shapefiles/MNCounties_MNDOT.shp", quiet = TRUE) %>%
  ms_simplify(keep = .01, keep_shapes = TRUE) %>%
  rename(countyfp = FIPS_CODE)

```

```{r master original}
master.original <- read.csv("Data/SLEDS/Masters/Master.csv") %>%
  mutate(n.institutions = ifelse(is.na(n.institutions), 0, n.institutions),
         ps.attended.in.same.ruca = ifelse(is.na(ps.attended.in.same.ruca), "Never attended ps", ps.attended.in.same.ruca),
         ps.attended.in.same.edr = ifelse(is.na(ps.attended.in.same.edr), "Never attended ps", ps.attended.in.same.edr),
         ps.attended.in.same.pr = ifelse(is.na(ps.attended.in.same.pr), "Never attended ps", ps.attended.in.same.pr),
         ps.attended.in.MN = ifelse(is.na(ps.attended.in.MN), "Never attended ps", ps.attended.in.MN),
         ps.grad = ifelse(ps.grad == "Y", 1, 0),
         ps.grad.InstitutionSector = ifelse(attended.ps == "No", "Never attended ps", ps.grad.InstitutionSector),
         ps.grad.InstitutionSector = ifelse(attended.ps == "Yes" & ps.grad == 0, "Did not grad", ps.grad.InstitutionSector),
         ps.grad.in.same.ruca = ifelse(attended.ps == "No", "Never attended ps", ps.grad.in.same.ruca),
         ps.grad.in.same.ruca = ifelse(attended.ps == "Yes" & ps.grad == 0, "Did not grad", ps.grad.in.same.ruca),
         ps.grad.in.same.edr = ifelse(attended.ps == "No", "Never attended ps", ps.grad.in.same.edr),
         ps.grad.in.same.edr = ifelse(attended.ps == "Yes" & ps.grad == 0, "Did not grad", ps.grad.in.same.edr),
         ps.grad.in.same.pr = ifelse(attended.ps == "No", "Never attended ps", ps.grad.in.same.pr),
         ps.grad.in.same.pr = ifelse(attended.ps == "Yes" & ps.grad == 0, "Did not grad", ps.grad.in.same.pr),
         ps.grad.in.MN = ifelse(attended.ps == "No", "Never attended ps", ps.grad.in.MN),
         ps.grad.in.MN = ifelse(attended.ps == "Yes" & ps.grad == 0, "Did not grad", ps.grad.in.MN),
         took.ACT = ifelse(is.na(ACTCompositeScore), "No", "Yes"),
         hs.grad.year = ifelse(hs.grad.year == 2008, 1,
                               ifelse(hs.grad.year == 2009, 2,
                                      ifelse(hs.grad.year == 2010, 3,
                                             ifelse(hs.grad.year == 2011, 4,
                                                    ifelse(hs.grad.year == 2012, 5,
                                                           ifelse(hs.grad.year == 2013, 6,
                                                                  ifelse(hs.grad.year == 2014, 7,
                                                                         ifelse(hs.grad.year == 2015, 8,
                                                                                ifelse(hs.grad.year == 2016, 9,
                                                                                       ifelse(hs.grad.year == 2017, 10,
                                                                                              ifelse(hs.grad.year == 2018, 11, 12))))))))))),
         attended.ps.within.first.year.hsgrad = ifelse(is.na(attended.ps.within.first.year.hsgrad), "Never attended PS", attended.ps.within.first.year.hsgrad),
         attended.ps.within.first.year.hsgrad = ifelse(attended.ps.within.first.year.hsgrad == "Yes", "Attended first year after HS",
                                                       ifelse(attended.ps == "No", "Never attended PS", "Attended but not first year"))) %>%
  drop_na(Dem_Desc)


write_csv(master.original, "Data/SLEDS/Masters/Master-updated-1.csv")

```

<br>

The purpose of multiple correspondence analysis is to examine any patterns among the independent variables with the goal of eliminating variables that are redundant - essentially two variables that are explaining the same thing. This type of analysis is not meant to confirm any predictions, but rather explore the data and make decisions on what data should be included in the prediction model.



# Independent variables

The master dataset is composed of `r comma(nrow(master.original))` rows. Each row represents an individual that graduated (StatusEnd = 8 or 9) from a high school located in EDR 6W, EDR 6E, and EDR 8 between 2008 and 2019.

The master dataset is also composed of `r comma(ncol(master.original))` columns. Columns mainly represent various socio-economic, demographic, and educational career characteristics of the indivdual as well as a few characteristics of the high school geographic from which they graduated (i.e. county unemployment). The variables for each individual is composed of a mix of categorical and numerical values. In this analysis, we will only be exploring categorical data.

Here is a list of all the categorical variables with a description of each.

```{r primary independent variables}
qual.var.names <-  c("PersonID", "hs.grad.year", "Gender", "LimitedEnglishProficiencyIndicator", "HomelessIndicator", "economic.status", "pseo.participant", "SpecialEdStatus", "non.english.home", "RaceEthnicity", "Dem_Desc", "edr", "took.ACT", "ap.exam", "cte.achievement", "english.learner", "sat.taken", "attended.ps.within.first.year.hsgrad", "attended.ps", "ps.grad", "ps.grad.InstitutionSector", "highest.cred.level")

quan.var.names <- c("total.cte.courses.taken", "cte.0", "cte.1", "cte.2", "cte.3", "cte.4", "cte.5", "cte.6", "cte.7", "cte.8", "cte.9", "cte.10", "cte.11", "cte.12", "cte.14", "cte.21", "cte.22", "cte.23", "cte.24", "cte.25", "cte.26", "avg.cte.intensity", "MCA.M", "MCA.R", "MCA.S", "avg.unemp.rate", "wages.3year.avg", "avg.wages.pct.state")

independent.var.names <- c("hs.grad.year", "Gender", "LimitedEnglishProficiencyIndicator", "HomelessIndicator", "economic.status", "pseo.participant", "SpecialEdStatus", "non.english.home", "RaceEthnicity", "Dem_Desc", "edr", "took.ACT", "ap.exam", "total.cte.courses.taken", "cte.achievement", "avg.cte.intensity", "MCA.M", "MCA.R", "MCA.S", "sat.taken", "attended.ps", "ps.grad", "highest.cred.level", "avg.unemp.rate", "wages.3year.avg")


cte.cc.var.names <- c("cte.1", "cte.2", "cte.3", "cte.4", "cte.5", "cte.6", "cte.7", "cte.8", "cte.9", "cte.10", "cte.11", "cte.12", "cte.14", "cte.21", "cte.22", "cte.23", "cte.24", "cte.25")

ind.qual.var <- master.original %>%
  select(all_of(`qual.var.names`)) %>%
  mutate_at(2:22, as.factor)



```

* **Demographics**
    + *Grad year:* this is a value representing each year starting with 1 representing 2008 and 12 representing 2019.
    + *Gender:* 
      + M = male
      + F = female
    + *RaceEthnicity:* 
      + AI = American Indian
      + Asian/PI = Asian or Pacific Islander
      + Black = African American
      + Hispanic = any race any Hispanic or Latino
      + Unknown = unknown race or ethnicity
      + White = Caucuasian and not hispanic

<br>
    
* **High school characteristics**
    + *Dem_Desc:* the RUCA category of the high school from which the individual graduated.
    + *edr:* the EDR of the high school from which the individual graduated.

<br>

* **High school enrollment**
    + *LimitedEnglishProficiencyIndicator:* 
      + Y = identified as having limited english proficiency at some point between 10th and 12th grade, otherwise N.
    + *HomelessIndicator:* 
      + Y = the individual was  identified as homeless at any point between 10th and 12th grade, otherwise N.
    + *economic.status:* 
      + 1 = the individual was eligible for free or reduced lunch (codes 1,2,4,5) at any point between 10th and 12th grade, otherwise 0.
    + *pseo.participant:* 
      + 1 = the individual participated in a PSEO course sometime between 10th and 12th grade, otherwise 0.
    + *SpecialEdStatus:* 
      + 1 = the individual required special education services at some point between 10th and 12th grade, otherwise 0.
    + *non.english.home:*   
      + 1 = the individual was identifed as having English as not the primary langauge spoken at home between 10th and 12th grad, otherwise 0.
    + *english.learner:* 
      + 1 = the individual was identified as an "english learner" at least one time between 10th and 12th grade, otherwise 0.
      
<br>
      
* **High school accomplishments**
    + *took.ACT:* whether an indiviual took the ACT exam
      + Yes
      + No
    + *ap.exam:*
      + 1 = the individual took an AP exam at some point, 
      + 0 = the individual did not take an AP exam.
    + *cte.achievement:* three indicators with 
      + "CTE concentrator or completor" being one, 
      + "CTE participant" meaning they took a CTE course but was not a concentrator or completor, and, 
      + "No CTE" meaning they didn't take a CTE course
    + *sat.taken:* 
      + 1 = individual took the SAT at some point between 10th and 12th grade, otherwise 0.

<br>

* **Post-secondary**
    + *attended.ps.within.first.year.hsgrad:* the individual attended a post-secondary institution within the first year of graduating high school.
      + Attended PS within first year
      + Atteded PS but not within first year
      + Never attended PS
    + *attended.ps:* 
      + Yes = the individual attended a post-secondary education institution, otherwise No.
    + *ps.grad:* whether the indiviual graduated from a post-secondary institution
    + *ps.grad.InstitutionSector:* the institution sector of the post-secondary school from which an individual has graduated.
      + 1 - Public, 4-year or above
      + 2 - Private not-for-profit, 4-year or above
      + 3 - Private for-profit, 4-year or above
      + 4 - Public, 2-year
      + 5 - Private not-for-profit, 2-year
      + 6 - Private for-profit, 2-year
      + 7 - Public, less-than 2-year
      + 8 - Private not-for-profit, less-than 2-year
      + 9 - Private for-profit, less-than 2-year
      + 10 - multiple sectors (if attended multiple institutions)
      + 99 - Sector unknown (not active)
      + "Never attended ps" = the individual never attended a post-secondary institution
      + "Did not grad" = the individual attended a post-secondary institution but did not graduate.
    + *highest.cred.level:* an identifier of the highest credential earned by the PersonID. The categories are;
      + Less than associate degree
      + Associate degree
      + Bachelors degree
      + Higher than bachelors degree.
    
<br>

# Frequencies

First, lets check to see there aren't any variable categories with very low frequencies which can distort the analysis.

The tables below shows there are a few to be concerned about.

* HomelessIndicator: There are only 161 individuals in the dataset that were confirmed homeless compared to 37,993 individuals that were never confirmed homeless.
* LimitedEnglishProficiency: There are only 805 individuals that were ever confirmed as having limited english proficiency compared to 37,349.
* grad.InstitutionSector: there very few individuals that graduated from a post secondary institution categorized as 5, 6, 7, 8 and 9. 
* RaceEthnicity: ther are only a 399 students categorized as American Indian and only 40 categorized as Unknown. 
* SAT.taken: Only 237 students who took the ACT.


<br>

```{r ind qual frequencies}
ind.qual.freq <- ind.qual.var %>%
  gather(key = "key", value = "value", 2:22) %>%
  group_by(key, value) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(data_id = seq(n()))

ind.qual.freq.plot <- ggplot(ind.qual.freq, aes(value, n)) +
  facet_wrap(~key, ncol = 4, scales = "free") +
  geom_col_interactive(position = "dodge", aes(data_id = data_id, tooltip = paste(key, "\n", value, "\nN = ", comma(n, accuracy = 1), sep = ""))) +
  labs(x="", y = "", color="", title = "N for each independent, categorical variable")+
  scale_y_continuous(labels=scales::comma)+
  theme_bar+
  scale_fill_manual(values = brewer.pal(n = 5, "RdYlBu"),
                    guide = guide_legend(ncol = 3)) +
  theme(legend.position = "bottom",
        text = element_text(size = 18),
        axis.text.x = element_text(angle = 25, hjust = .9))


girafe(ggobj = ind.qual.freq.plot, width_svg = 10, height_svg = 20) %>%
  girafe_options(opts_selection(type = "none"),
                 opts_sizing(rescale = FALSE))      

```

<br>

Due to these low frequencies, I'm going to do the following;

* HomelessIndicator: eliminate from analysis
* LimitedEnglishProficiency: Keep for now
* grad.InstitutionSector: combine categories 5,6,7,8 and 9 into one category. This means that the following institution sectors will be relabeled as "11";
    + 5 - Private not-for-profit, 2-year
    + 6 - Private for-profit, 2-year
    + 7 - Public, less-than 2-year
    + 8 - Private not-for-profit, less-than 2-year
    + 9 - Private for-profit, less-than 2-year
* RaceEthnicity: Keep for now
* SAT.taken: eliminate from analysis.

```{r updated ind qual var}
master.ind.qual.var <- ind.qual.var %>%
  select(-PersonID, -HomelessIndicator, -sat.taken) %>%
  mutate(ps.grad.InstitutionSector = ifelse(ps.grad.InstitutionSector %in% c("5", "6", "7", "8", "9"), "11", as.character(ps.grad.InstitutionSector)),
         ps.grad.InstitutionSector = as.factor(ps.grad.InstitutionSector))

master.ind.qua.var.write <- ind.qual.var %>%
  select(-HomelessIndicator, -sat.taken) %>%
  mutate(ps.grad.InstitutionSector = ifelse(ps.grad.InstitutionSector %in% c("5", "6", "7", "8", "9"), "11", as.character(ps.grad.InstitutionSector)),
         ps.grad.InstitutionSector = as.factor(ps.grad.InstitutionSector))

```

<br>

# Demographics

<br>

Below are the variables that are categorized as "demographics". When I initially did this analysis, the category "Unknown" in the RaceEthnicity variable was throwing off the analysis. So we will filter out those individuals for which there are only `r comma(nrow(filter(master.ind.qual.var, RaceEthnicity == "Unknown")))` individuals.

```{r demographics prep, echo=TRUE}
demographics <- master.ind.qual.var %>%
  select(Gender, RaceEthnicity) %>%
  filter(RaceEthnicity != "Unknown")

head(demographics)

names(demographics)
```

<br>

After filtering out the "Unknown" RaceEthnicity category, we have `r comma(nrow(demographics))` individuals in the dataset.


## MCA analysis

The charts below show that the variable RaceEthnicity and hs.grad.year may have significantly more differentiation among them than Gender.


```{r MCA, echo=TRUE}
demographics.mca <- MCA(demographics, graph = TRUE)
```

<br>


## Eigenvalues

The eigenvalues measure the amount of variation retained by each principle component. This will help us determine how many principle components to be considered.

The values below indicate that the first dimension explains nearly 21% of the variation with each subsequent dimension explaining an additional 20%. It takes only 4 dimensions to explain over 75%.

```{r demographics eigenvalues}
eig.val <- get_eigenvalue(demographics.mca)

eig.val
```

<br>

## Correlation between variables and principle dimensions

Let's take a look to see which variables are going to be important. The chart below shows that the variables related to the graduation years and race/ethnicity are highly correlated with the first dimension while the Gender variable is correlated with dimension 2. 

```{r demogrpahics mca correlation between variables}
fviz_mca_var(demographics.mca, 
             choice = "mca.cor",
             ggtheme = theme_minimal())

```

<br>

## Quality of representation

From the chart below, we can see that the variables related to race and ethnicity have categories that are very different from each other - in particular black and Asian/PI. White and American Indian don'e differentiate themselves very much. As expected, Male and Female are opposite of each other.


```{r demographics quality of representation}
fviz_mca_var(demographics.mca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, # Avoid text overlapping
             ggtheme = theme_minimal())

```

<br>

Let's turn this into a bar plot to better visualize all of these. Since we know, according to the eigenvalues, that the first few dimensions only represent 20% of the variance, and each subsequent dimension only explains a bit more variance, it's likely that it will take numerous dimensions to explain a sufficient amount of our variance. Therefore, we will graph visualize the quality of representation up to the max number of dimensions allowed - 5. 

This shows that the RaceEthnicity categories and the gender categories have the highest Cos2.


```{r visualize quality of representation}
fviz_cos2(demographics.mca, choice = "var", axes = 1:2)

```

<br>

For demographics, it seems that RaceEthnicity and Gender are worth keeping around.

<br>

# High school characteristics

<br>

Below are the variables that are categorized as "high school characteristics". 

```{r hs characteristics prep, echo=TRUE}
hs.character <- master.ind.qual.var %>%
  select(Dem_Desc, edr) 

head(hs.character)

names(hs.character)
```

<br>


## MCA analysis

The charts below show that the Economic Development Region the high school is located in and the RUCA category of the high school county may have significant differentiation among them.


```{r hs character MCA, echo=TRUE}
hs.character.mca <- MCA(hs.character, graph = TRUE)
```

<br>


## Eigenvalues

The eigenvalues measure the amount of variation retained by each principle component. This will help us determine how many principle components to be considered.

The values below indicate that the first dimension explains nearly 34% of the variation and the second dimension explaining another nearl 30%. It only takes 3 dimensions to rise above the 75% threshold.

```{r hs character eigenvalues}
eig.val <- get_eigenvalue(hs.character.mca)

eig.val
```

<br>

## Correlation between variables and principle dimensions

Let's take a look to see which variables are going to be important. The chart below shows that the variables related to the EDR and and RUCA category of the high school are related to both dimensions equally. 

```{r hs character mca correlation between variables}
fviz_mca_var(hs.character.mca, 
             choice = "mca.cor",
             ggtheme = theme_minimal())

```

<br>

## Quality of representation

So far, we've only looked at the first two dimensions which only explain 15% of the variance. The quality of representation will help us assess the degree of association between variable categories and a particular axis. If the variable is well represented by two dimensions, the sum of COS2 is close to one. For some of the row items, more than 2 dimensions are required to perfectly represent the data.

From the table below, we can see that the variables make a lot of sense in terms of how they are related. The entirely rural Dem_Desc category is very much related to the EDR that is most rural - EDR 6W. The more metro based EDR - EDR 6E - is more closely aligned with Urban/town/rural mix RUCA categories and EDR 8 is more closely aligned with the town/rural mix RUCA category. This indicates that we could likely use just one of the variables.


```{r hs character quality of representation}
fviz_mca_var(hs.character.mca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, # Avoid text overlapping
             ggtheme = theme_minimal())

```

<br>

Since it takes 3 dimensions to reach the 75% threshold, the chart below provides the Cos2 quality of representation for those three dimensions. It's close, but this shows that the RUCA categories might be a bit more telling, especially with those first two dimensions. But it's close and I think using either the Dem_Desc or the EDR alone would be fine.


```{r hs character visualize quality of representation}
fviz_cos2(hs.character.mca, choice = "var", axes = 1:3)

```

<br>

For high school characteristics, it seems that using just the EDR variable might be enough.

<br>

# High school enrollment

<br>

Below are the variables that are categorized as "high school enrollment". 

```{r hs enroll prep, echo=TRUE}
hs.enroll <- master.ind.qual.var %>%
  select(LimitedEnglishProficiencyIndicator, economic.status, pseo.participant, SpecialEdStatus, non.english.home, english.learner) 

head(hs.enroll)

names(hs.enroll)

```

<br>


## MCA analysis

The charts below show that the one grouping related to english speaking proficiency that might be closely aligned to each other - the variables LimitedEnglishProficiencyIndicator, non.english.home, and english learner.


```{r hs enroll MCA, echo=TRUE}
hs.enroll.mca <- MCA(hs.enroll, graph = TRUE)
```

<br>


## Eigenvalues

The eigenvalues measure the amount of variation retained by each principle component. This will help us determine how many principle components to be considered.

The values below indicate that the first dimension explains nearly 37% of the variation and the second dimension explaining another 20%. It only takes 4 dimensions to rise above the 75% threshold.

```{r hs enroll eigenvalues}
eig.val <- get_eigenvalue(hs.enroll.mca)

eig.val
```

<br>

## Correlation between variables and principle dimensions

Let's take a look to see which variables are going to be important. The chart below shows that the variables related to english speaking and economic status are more related to dimension 1 while pseo participantion and special ed status are more aligned with the second dimension.

```{r hs enroll mca correlation between variables}
fviz_mca_var(hs.enroll.mca, 
             choice = "mca.cor",
             ggtheme = theme_minimal())

```

<br>

## Quality of representation

So far, we've only looked at the first two dimensions which only explain 15% of the variance. The quality of representation will help us assess the degree of association between variable categories and a particular axis. If the variable is well represented by two dimensions, the sum of COS2 is close to one. For some of the row items, more than 2 dimensions are required to perfectly represent the data.

From the chart below, we can see that the variables make a lot of sense in terms of how they are related. The higheset Cos2 are all related to the english proficiency variable. Economic status has a relativly lower Cos2. Special education status was also a bit higher.


```{r hs enroll quality of representation}
fviz_mca_var(hs.enroll.mca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, # Avoid text overlapping
             ggtheme = theme_minimal())

```

<br>

From the english proficiency group, it seems that the LimitedEnglishProficencyIndicator might be the best to use in terms of that group, while it is good to probably keep Economic status and special ed status.


```{r hs enroll visualize quality of representation}
fviz_cos2(hs.enroll.mca, choice = "var", axes = 1:3)

```

<br>

For high school enrollment, it seems that using the following variables would be best;

* LimitedEnglishProficiencyIndicator
* SpecialEdStatus
* PSEO Participant
* Economic status

<br>

# High school accomplishments

<br>

Below are the variables that are categorized as "high school accomplishments". 

```{r hs accom prep, echo=TRUE}
hs.accom <- master.ind.qual.var %>%
  select(took.ACT, ap.exam, cte.achievement) 

head(hs.accom)

names(hs.accom)

```

<br>


## MCA analysis

The charts below show that the one grouping related to whether the ACT and ap were taken might be closely aligned to each other.


```{r hs accom MCA, echo=TRUE}
hs.accom.mca <- MCA(hs.accom, graph = TRUE)
```

<br>


## Eigenvalues

The eigenvalues measure the amount of variation retained by each principle component. This will help us determine how many principle components to be considered.

The values below indicate that the first dimension explains nearly 33% of the variation and the second dimension explaining another 25%. It only takes 3 dimensions to rise above the 75% threshold.

```{r hs accom eigenvalues}
eig.val <- get_eigenvalue(hs.accom.mca)

eig.val
```

<br>

## Correlation between variables and principle dimensions

Let's take a look to see which variables are going to be important. The chart below shows that the variables related to english speaking and economic status are more related to dimension 1 while pseo participantion and special ed status are more aligned with the second dimension.

```{r hs accom mca correlation between variables}
fviz_mca_var(hs.accom.mca, 
             choice = "mca.cor",
             ggtheme = theme_minimal())

```

<br>

## Quality of representation

So far, we've only looked at the first two dimensions which only explain 15% of the variance. The quality of representation will help us assess the degree of association between variable categories and a particular axis. If the variable is well represented by two dimensions, the sum of COS2 is close to one. For some of the row items, more than 2 dimensions are required to perfectly represent the data.

From the chart below, we can see that the variables make a lot of sense in terms of how they are related. The higheset Cos2 are all related to the CTE achievement variable, and they are spread very far apart meaning lots of differentiation there. The CTE Concentrator or completor was also much more aligned with not taking the ACT or AP exam. Not suprisingly, the designation on whether an individual took the AP exam and the ACT are closely aligned with individuals that took the ACT were more aligned with individuals that also took the AP exam. However, the ACT variable is much stronger.

```{r hs accom quality of representation}
fviz_mca_var(hs.accom.mca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, # Avoid text overlapping
             ggtheme = theme_minimal())

```

<br>

Judging by the Cos2 values, it seems that using the ACT variable instaed of the AP exam would be just fine.


```{r hs accom visualize quality of representation}
fviz_cos2(hs.accom.mca, choice = "var", axes = 1:3)

```

<br>

For high school accomplishments, it seems that using the following variables would be best;

* cte.achievement
* took.ACT

<br>

# Post-secondary

<br>

Below are the variables that are categorized as "post-secondary". 

```{r post secondary prep, echo=TRUE}
ps <- master.ind.qual.var %>%
  select(attended.ps.within.first.year.hsgrad, attended.ps, ps.grad, ps.grad.InstitutionSector, highest.cred.level) 

head(ps)

names(ps)

```

<br>


## MCA analysis

The charts below show that the following variables form a bit of a group - attended.ps.within.first.year.hsgrad, ps.grad, attended.ps.

```{r ps MCA, echo=TRUE}
ps.mca <- MCA(ps, graph = TRUE)
```

<br>


## Eigenvalues

The eigenvalues measure the amount of variation retained by each principle component. This will help us determine how many principle components to be considered.

The values below indicate that the first dimension explains nearly 27% of the variation and the second dimension explaining another 14%. It only takes 6 dimensions to rise above the 75% threshold.

```{r ps eigenvalues}
eig.val <- get_eigenvalue(ps.mca)

eig.val
```

<br>

## Correlation between variables and principle dimensions

Let's take a look to see which variables are going to be important. The chart below shows that the variables related t whather they attended post-seconary form a group.

```{r ps mca correlation between variables}
fviz_mca_var(ps.mca, 
             choice = "mca.cor",
             ggtheme = theme_minimal())

```

<br>

## Quality of representation

So far, we've only looked at the first two dimensions which only explain 15% of the variance. The quality of representation will help us assess the degree of association between variable categories and a particular axis. If the variable is well represented by two dimensions, the sum of COS2 is close to one. For some of the row items, more than 2 dimensions are required to perfectly represent the data.

From the chart below, we can see that the variables make a lot of sense in terms of how they are related. The categories indicating whether someone attended post-secondary are all in the same spot - top left corner. The categories indicating whether they graduated post-secondary are all group together in the bottom left corner.

```{r ps quality of representation}
fviz_mca_var(ps.mca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, # Avoid text overlapping
             ggtheme = theme_minimal())

```

<br>

Judging by the Cos2 values, it seems that using the ps.grad.InstitutionSector variable will do a fine job at covering attending post-secondary and atteded within first year of high school graduation. 


```{r ps visualize quality of representation}
fviz_cos2(ps.mca, choice = "var", axes = 1:5)

```

<br>

For post-secondary, it seems that using the following variables would be best;

* ps.grad.InstitutionSector
* highest.cred.level

<br>

# Variables to consider keeping and eliminating

The analysis above shows that the following variables would likely provide high quality representation without muddying up the analysis with so many variables.

*Demographics
  * Gender
  * RaceEthnicity
* High school characteristics
  * EDR
* High school enrollment
  * LimitedEnglishProficiency
  * SpecialEdStatus
  * Economic.status
  * PSEO participant
* High school accomplishments
  * cte.achievement
  * took.act
* Post-secondary
  * ps.grad.InstitutionSector
  * highest.cred.level
  
  
This means we will eliminate the following variables

* High school characteristics: Dem_Desc
* High school enrollment: non.english.home & english.learner
* High school accomplishments: ap.exam
* Post-secondary: attended.ps.within.first.year.hsgrad & attended.ps & ps.grad

```{r final master}
master.final.after.mca <- master.ind.qua.var.write %>%
  select(-hs.grad.year, -Dem_Desc, -non.english.home, -english.learner, -ap.exam, -attended.ps.within.first.year.hsgrad, -attended.ps, -ps.grad)

write_csv(master.final.after.mca, "Data/SLEDS/Masters/Master-final-after-mca.csv")

```
