---
title: "Analysis - local employment CART"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(readxl)
library(janitor)
library(lubridate)
library(systemfonts)
reset_font_cache()
library(ggtext)
library(gmodels)
library(fastDummies)
library(car)
library(glmnet)
library(glmnetUtils)
library(pscl)
library(sjPlot)
library(rpart)
library(rpart.plot)
```

```{r join docs, include=FALSE}
theme_bar <- theme_bw() +
  theme(panel.grid.major = element_line(color = "grey70", size = 0.1),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.key.size = unit(1, "lines"),
        legend.margin = margin(0,0,0,0),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        text = element_text(family = "Arial") ,
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"))

theme_line <- theme_bw() +
  theme(legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.text = element_text(margin = margin(l = 2)),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey70", size = 0.1),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial") ,
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"))


theme_sf <- theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "white"),
        panel.border = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial") ,
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"))

regions <- read_csv("Data/Join docs/county_regions.csv") %>%
    select(5,6) %>%
    unique() %>%
    mutate(edr = str_replace(edr, "  ", " "),
           planning.region = str_replace(planning.region, " Minnesota", ""),
           planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
           edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"))

counties.regions <- read_csv("Data/Join docs/county_regions.csv") %>%
  rename(mif = `MIF Region`) %>%
  mutate(countyfp = formatC(countyfp, width = 3, flag = "0"),
         Name = str_to_title(Name),
         Name = str_replace(Name, "Q", "q"),
         Name = str_replace(Name, "Of The", "of the"),
         Name = str_replace(Name, "Mcleod", "McLeod"),
         Dem_Desc = ifelse(Name == "Minnesota", "Minnesota", Dem_Desc) ,
         edr = str_replace(edr, "  ", " "),
         planning.region = str_replace(planning.region, " Minnesota", ""),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"),
         mif = ifelse(is.na(mif), "TC", mif),
         mif = as.factor(mif),
         mif = fct_relevel(mif, "NW", "NE", "WC", "EC", "SW", "SE", "TC"))


color.ruca <- c("Entirely rural" = "#009933", "Town/rural mix" = "#99CC33", "Urban/town/rural mix" = "#CC9966", "Entirely urban" = "#754C29", "Minnesota" = "black")

color.pr <- c("Northwest" = 	"#4575b4", "Northeast" = "grey", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black")

color.edr <- c("EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365", "Minnesota" = "black")

color.pr.edr <- c ("Northwest" = "#4575b4","Northeast" = "#e0f3f8", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black", "EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365")

mn_counties <- st_read("Data/Shapefiles/county shapefiles/MNCounties_MNDOT.shp", quiet = TRUE) %>%
  ms_simplify(keep = .01, keep_shapes = TRUE) %>%
  rename(countyfp = FIPS_CODE)


```

```{r master dataset}
master.after.ct <- read_csv("Data/SLEDS/Masters/Master-after-ct.csv") %>%
    mutate_at(c(2:16, 18:23, 25, 29:32), as.factor) 

```


<br>

We will use a tree-based method for classification - CART analysis. This type of analysis involves stratifying and/or segmenting the predictor space into a number of simple regions. Essentially, it's another way to see which independent variables play a role in if an individual has meaningful employment in the local region X. We will be using the independent variables that were identified as being important in the multiple correspondence analysis.

There are three "times" in which we checked to see if the individual had meaningful employment within these geographies;

1. the same year as graduation
2. two years after graduation, and
3. seven years after graduation. 

Meaningful employment is determined by whether an individual worked 1,000 hours for an employer during time X. In addition, if the individual worked 1,000 hours for that employer during another year, but not for that exact "time X" year, it's still considered "meaningful". 

Due to time x potentially being passed the date of the latest data (2019) for some individuals, the analysis below will filter out all individuals where time x is after 2019. 

The primary independent variables that the cross tables indicated as important are;

```{r list of ind var}
ind.var <- master.after.ct %>%
  select(-PersonID, -hs.grad.year.county.match, -hs.grad.year.2.county.match, -hs.grad.year.7.county.match, -hs.grad.year.edr.match, -hs.grad.year.2.edr.match, -hs.grad.year.7.edr.match, -hs.grad.year.region.match, -hs.grad.year.region.match, -hs.grad.year.2.region.match, -hs.grad.year.7.region.match, -hs.grad.year.state.match, -hs.grad.year.2.state.match, -hs.grad.year.7.state.match) 

dep.var.names <- master.after.ct %>%
  select(hs.grad.year.county.match, hs.grad.year.2.county.match, hs.grad.year.7.county.match, hs.grad.year.edr.match, hs.grad.year.2.edr.match, hs.grad.year.7.edr.match, hs.grad.year.region.match, hs.grad.year.region.match, hs.grad.year.2.region.match, hs.grad.year.7.region.match, hs.grad.year.state.match, hs.grad.year.2.state.match, hs.grad.year.7.state.match) %>%
  names()

ind.var.names <- ind.var %>%
  names()

ind.var %>%
  lapply(class)
```
However, there are plenty of variables here that are VERY similar. For example, whether an individual graduated from a post-secondary institution and the type of institution they graduated from. So let's simplify the variables to what is most important and has the least amount of overlap with other variables.

* Demographic
    + RaceEthnicity
* High school characteristics
    + Dem_Desc
    + EDR
    + avg.wages.pct.state
* High school enrollment
    + LimitedEnglishProficiencyIndicator
    + PSEO.participant
* High school accomplishments
    + took.act
    + ap.exam
    + cte.achievement
    + MCA.M
    + MCA.R
    + MCA.S
* Post-secondary
    + ps.grad.InstitutionSector
    + highest.cred.level
    
We eliminated the following variables for now;

* non.english.home
* english.learner
* total.cte.courses.taken
* ps.grad
* attended.ps

In addition to eliminating those variables, we will also simplify the ps.grad.InstitutionSector to the following codes;

* 1 = graduated from the following insitutions
    + Public, 4-year (1)
    + Private, not for profit 4-year (2)
    + Private, for profit 4-year (3)
* 2 = graduated from the following institutions
    + Private, not for profit 2-year (5, 11)
    + Private, for profit 2-year (6, 11)
    + Public, less than 2-year (7, 11)
    + Private, not for profit less than 2-year (8, 11)
    + Private for-profit, less-than 2-year (9, 11)
* 3 = graduated from multiple institution sectors (10)
* 4 = graduated from a public, 2-year institution (4,4)
* 5 = Did not graduate or attend a post-secondary institution

```{r updating master}
updated.master <- master.after.ct %>%
  select(`dep.var.names`, RaceEthnicity, Dem_Desc, edr, avg.wages.pct.state, LimitedEnglishProficiencyIndicator, pseo.participant, took.ACT, ap.exam, cte.achievement, MCA.M, MCA.R, MCA.S, ps.grad.InstitutionSector, highest.cred.level) %>%
  mutate(ps.grad.InstitutionSector = ifelse(ps.grad.InstitutionSector %in% c(1,2,3), 1,
                                            ifelse(ps.grad.InstitutionSector == 11, 2,
                                                   ifelse(ps.grad.InstitutionSector == 10, 3,
                                                          ifelse(ps.grad.InstitutionSector == 4, 4, 5))))) 

updated.ind.var.names <- updated.master %>%
  select(-`dep.var.names`) %>%
  names()

```

<br>

# Steps in analysis

Our dependent variable - the individual has meaningful employment in same location X as their high school - has 4 different outcomes.

1. Location match: the individual has
    + a MN employment record at time X
    + has meaningful employment at time X
    + is employed in the same location X as their high school.
2. No location match: the individual has 
    + a MN employment record at time X
    + has meaningful employment at time X
    + does NOT have employment in the same location X as their high school.
3. No meaningful emp: the individual has
    + a MN employment record at time X
    + the employment is not meaningful at time X.
4. No MN emp record: the individual has
    + NO MN employment record at time X.
    
Since we are most interested in whether an individual has meaningful employment in region X at time X we will create two groups; 1. all individuals that meet the criteria/outcome 1 from list above, and 2. all remaining individuals.
    
For each time X we will analyze whether the independent variables are good at predicting whether an individual has a MN employment record AND has meaningful employment AND the meaningful employment is in the same region X as their high school.



<br>

# Analysis - time X = grad year +0

## County meaningful employment

```{r master CART grad year county}
master.grad.year.county.path <- updated.master %>%
  select(hs.grad.year.county.match, `updated.ind.var.names`)  %>%
  filter(hs.grad.year.county.match != "Unknown") %>%
  filter(hs.grad.year.county.match != "After 2019") %>%
  drop_na(hs.grad.year.county.match) %>%
  mutate(hs.grad.year.county.match = ifelse(hs.grad.year.county.match == "Location match", "Match", "No match")) 
  
```

This is an overly complex tree.

<br>

```{r cart county emp grad year plus zero, echo=TRUE}
model <- rpart(hs.grad.year.county.match ~ ., data = master.grad.year.county.path, method = "class", cp = .000252)

model
rpart.plot(model)
```

<br>

Test dataset.

