---
title: "AP"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(extrafont)
library(readxl)
library(janitor)
library(lubridate)

loadfonts()

```

<br>

# Data prep


The next dataset to merge with the master is the enrollment in AP courses. 

<br>

```{r ap original data}
ap.original <- read_csv("Data/SLEDS/AP/AP.csv") %>%
  mutate(PersonID = as.integer(PersonID)) %>%
  drop_na(PersonID)

head(ap.original)

names(ap.original)

```

<br>

The AP dataset has `r comma(nrow(ap.original), accuracy = 1)` observations and `r comma(ncol(ap.original), accuracy = 1)` columns. Each observation represents an AP exam taken so this means that a PersonID can be listed multiple times if the individual took multiple AP exams. The ExamGrade is on a 1 - 5 scale with the highest scores being better. According to Meredith at OHE, colleges give credit if the score is 3 or higher. 

I'm a bit torn here. There seems to be three approaches to how I can utilize this data in the model. 

1. I could either just create a category of whether an individual took an AP course and not care what the grade was.
2. I could only use individuals who scored 3 or higher.
3. I could keep all the data and change it so that each observation represents a PersonID and new columns created for each AP exam and their score. 

For now, I think I'm going to just utilize option 1 - whether they took an AP exam or not. 

<br>

```{r ap took exam}
ap.tookexam <- ap.original %>%
  distinct(PersonID) %>%
  mutate(ap.exam = 1)

head(ap.tookexam)

names(ap.tookexam)

```

<br>

The updated AP dataset shows that `r comma(nrow(ap.tookexam), accuracy = 1)` took at least one AP exam. Now it's time to merge it with the master list. For the PersonID that didn't take an AP exam, they will receive a value of 0 in that column.

<br>

```{r ap join with master}
master.3 <- master.2 %>%
  left_join(ap.tookexam, by = "PersonID") %>%
  mutate(ap.exam = ifelse(is.na(ap.exam), 0, ap.exam))

head(master.3)

names(master.3)
```

<br>

The updated master spreadsheet continues to have `r comma(nrow(master.3), accuracy = 1)` observations and now has `r comma(ncol(master.3), accuracy = 1)` columns.

<br>

# Summary of AP data