---
title: "AP"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(readxl)
library(janitor)
library(lubridate)
library(systemfonts)
reset_font_cache()
library(ggtext)


```

```{r themes and join docs, include=FALSE}
theme_sf <- theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "white"),
        panel.border = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))

mn_counties <- st_read("Data/Shapefiles/county shapefiles/MNCounties_MNDOT.shp", quiet = TRUE) %>%
  ms_simplify(keep = .01, keep_shapes = TRUE) %>%
  rename(countyfp = FIPS_CODE)

counties.regions <- read_csv("Data/Join docs/county_regions.csv") %>%
  rename(mif = `MIF Region`) %>%
  mutate(countyfp = formatC(countyfp, width = 3, flag = "0"),
         Name = str_to_title(Name),
         Name = str_replace(Name, "Q", "q"),
         Name = str_replace(Name, "Of The", "of the"),
         Name = str_replace(Name, "Mcleod", "McLeod"),
         Dem_Desc = ifelse(Name == "Minnesota", "Minnesota", Dem_Desc) ,
         edr = str_replace(edr, "  ", " "),
         planning.region = str_replace(planning.region, " Minnesota", ""),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"),
         mif = ifelse(is.na(mif), "TC", mif),
         mif = as.factor(mif),
         mif = fct_relevel(mif, "NW", "NE", "WC", "EC", "SW", "SE", "TC"))

```

<br>

# Data prep


The next dataset to merge with the master is the enrollment in AP courses. 

<br>

```{r ap original data}
ap.original <- read_csv("Data/SLEDS/AP/AP.csv") %>%
  mutate(PersonID = as.integer(PersonID)) %>%
  drop_na(PersonID)

head(ap.original)

names(ap.original)

```

<br>

The AP dataset has `r comma(nrow(ap.original), accuracy = 1)` observations and `r comma(ncol(ap.original), accuracy = 1)` columns. Each observation represents an AP exam taken so this means that a PersonID can be listed multiple times if the individual took multiple AP exams. The ExamGrade is on a 1 - 5 scale with the highest scores being better. According to Meredith at OHE, colleges give credit if the score is 3 or higher. 

I'm a bit torn here. There seems to be three approaches to how I can utilize this data in the model. 

1. I could either just create a category of whether an individual took an AP exam and not care what the grade was.
2. I could only use individuals who scored 3 or higher.
3. I could keep all the data and change it so that each observation represents a PersonID and new columns created for each AP exam and their score. 

For now, I think I'm going to just utilize option 1 - whether they took an AP exam or not. 

<br>

```{r ap took exam}
ap.tookexam <- ap.original %>%
  distinct(PersonID) %>%
  mutate(ap.exam = 1)

head(ap.tookexam)

names(ap.tookexam)

```

<br>

The updated AP dataset shows that `r comma(nrow(ap.tookexam), accuracy = 1)` took at least one AP exam. Now it's time to merge it with the master list. For the PersonID that didn't take an AP exam, they will receive a value of 0 in that column.

<br>

```{r ap join with master}
master.3 <- read_csv("Data/SLEDS/Masters/master-2.csv") %>%
  left_join(ap.tookexam, by = "PersonID") %>%
  mutate(ap.exam = ifelse(is.na(ap.exam), 0, ap.exam))

head(master.3)

names(master.3)
```

<br>

The updated master spreadsheet continues to have `r comma(nrow(master.3), accuracy = 1)` observations and now has `r comma(ncol(master.3), accuracy = 1)` columns.

<br>

# Summary of AP data

Similar to ACT, lets check on whether the rates of students taking an AP exam differ across regions, RUCA groups or counties.

<br>

```{r table ap summary total, echo=FALSE}
ap.summary.total <- master.3 %>%
  select(PersonID, ap.exam) %>%
  mutate(ap.exam = ifelse(ap.exam == 0, "Did not take AP exam", "Took AP exam")) %>%
  tabyl(ap.exam) %>%
  mutate(percent = scales::percent(percent, accuracy = .1))

datatable(ap.summary.total, class = "cell-border stripe", filter = "top", rownames = FALSE)
```
<br>

Out of the total observations in our dataset, 15% have taken an AP exam.

<br>

```{r table ap exam ruca, echo=FALSE}
ap.summary.ruca <- master.3 %>%
  select(Dem_Desc, ap.exam) %>%
  mutate(ap.exam = ifelse(ap.exam == 0, "Did not take AP exam", "Took AP exam")) %>%
  tabyl(Dem_Desc, ap.exam) %>%
  mutate(percent.took.exam = `Took AP exam` / (`Did not take AP exam` + `Took AP exam`),
         percent.took.exam = scales::percent(percent.took.exam, accuracy = .1))

datatable(ap.summary.ruca, class = "cell-border stripe", filter = "top", rownames = FALSE)
  
```
<br>

Interestingly, there was significant difference between the percentage of students who took an AP exam in entirely rural schools vs. the other RUCA categories. Only 2% of students that graduated from an entirely rural school took an AP exam compared to 17% and 20% in town/rural mix and urban/town/rural mix schools.

<br>

```{r table ap summary edr, echo=FALSE}
ap.summary.edr <- master.3 %>%
  select(edr, ap.exam) %>%
  mutate(ap.exam = ifelse(ap.exam == 0, "Did not take AP exam", "Took AP exam")) %>%
  tabyl(edr, ap.exam) %>%
  mutate(percent.took.exam = `Took AP exam` / (`Did not take AP exam` + `Took AP exam`),
         percent.took.exam = scales::percent(percent.took.exam, accuracy = .1))

datatable(ap.summary.edr, class = "cell-border stripe", filter = "top", rownames = FALSE)
```

<br>

EDR 6W had the lowest percentage of graduates in the dataset that took an AP exam with 8%. This was quite a bit lower than EDR 6E and 8 which had 18% and 16% of their students take one, respectively. This corresponds to the RUCA data since EDR 6W is going to have quite a bit more entirely rural schools.

<br>

```{r map ap exam county, echo=FALSE}
ap.summary.county <- master.3 %>%
  select(county.name, ap.exam) %>%
  mutate(ap.exam = ifelse(ap.exam == 0, "Did not take AP exam", "Took AP exam")) %>%
  tabyl(county.name, ap.exam) %>%
  mutate(percent.took.exam = `Took AP exam` / (`Did not take AP exam` + `Took AP exam`)) %>%
  left_join(counties.regions, by = c("county.name" = "Name")) %>%
  right_join(mn_counties[,c(4,7)], by = "countyfp")

ap.summary.county.map <- ggplot(ap.summary.county) +
  geom_sf_interactive(color = "grey85", aes(geometry = geometry, fill = percent.took.exam, data_id = countyfp, tooltip = paste(county.name, "\nNumber of students who took exam: ", comma(`Took AP exam`, accuracy = 1), "\nPercent of students who took AP exam: ", percent(percent.took.exam, accuracy = .1), sep = ""))) +
  theme_sf+
  scale_fill_fermenter(palette = "PuBu", direction = 1, labels = scales::percent) +
  labs(title = "Percent of students who took AP exam by county") +
  theme(legend.box.margin = margin(50, 0, 0, -100),
        text = element_text(size = 18))

girafe(ggobj = ap.summary.county.map, height_svg = 10, width_svg = 10) %>%
  girafe_options(opts_sizing(rescale = FALSE))

```

<br>

There is significant variation in the percent of students who've taken an AP exam across counties. It ranges from 0.1% in Murray County to 26.0% in Redwood County. 

<br>


```{r write master, include=FALSE}
write_csv(master.3, "Data/SLEDS/Masters/Master-3.csv")
```