---
title: "Assessments"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(readxl)
library(janitor)
library(lubridate)
library(systemfonts)
reset_font_cache()
library(ggtext)
library(gmodels)
```

```{r themes and join docs, include=FALSE}
theme_line <- theme_bw() +
  theme(legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.text = element_text(margin = margin(l = 2)),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey70", size = 0.1),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))

theme_sf <- theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "white"),
        panel.border = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))

theme_bar <- theme_bw() +
  theme(panel.grid.major = element_line(color = "grey70", size = 0.1),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.key.size = unit(1, "lines"),
        legend.margin = margin(0,0,0,0),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        text = element_text(family = "Arial"))


mn_counties <- st_read("Data/Shapefiles/county shapefiles/MNCounties_MNDOT.shp", quiet = TRUE) %>%
  ms_simplify(keep = .01, keep_shapes = TRUE) %>%
  rename(countyfp = FIPS_CODE)

counties.regions <- read_csv("Data/Join docs/county_regions.csv") %>%
  rename(mif = `MIF Region`) %>%
  mutate(countyfp = formatC(countyfp, width = 3, flag = "0"),
         Name = str_to_title(Name),
         Name = str_replace(Name, "Q", "q"),
         Name = str_replace(Name, "Of The", "of the"),
         Name = str_replace(Name, "Mcleod", "McLeod"),
         Dem_Desc = ifelse(Name == "Minnesota", "Minnesota", Dem_Desc) ,
         edr = str_replace(edr, "  ", " "),
         planning.region = str_replace(planning.region, " Minnesota", ""),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"),
         mif = ifelse(is.na(mif), "TC", mif),
         mif = as.factor(mif),
         mif = fct_relevel(mif, "NW", "NE", "WC", "EC", "SW", "SE", "TC"))

```

<br>

# Data prep

Okay, lets check out IPEDS.

<br>

```{r ipeds original}
assessments.original <- read_csv("Data/SLEDS/K12 Assessments/K12Assessment.csv")  %>%
  mutate(PersonID = as.integer(PersonID),
         TestName = as.factor(TestName),
         Subject = as.factor(Subject),
         AchievementLevel = as.factor(AchievementLevel)) 

head(assessments.original)

names(assessments.original)

```

<br>

We have `r comma(assessments.original %>% nrow(), accuracy = 1)` rows and `r comma(assessments.original %>% ncol(), accuracy = 1)` columns in this dataset. Here are the descriptions for each of the columns.

* K12AsstTimeID: Day, month, and year the assessment took place.
* TestName: Name of the test taken. There are 12 different tests in the dataset - 
  + ACCESS: English proficiency, 
  + ALTACCESS: English proficiency for special education or meet requirements
  + MCA-II & MCA-III: Minnesota Comprehsensive Assessments (series 2 and 3)
  + MOD-II, MOD-III: Modified Comprehensive Assessment, used only until 2014
  + MTAS & MTAS-III: Minnesota Test of Academic Skills
  + MTELL:
  + SOLOM: Minnesota Student Oral Language Observation Matrix
  + TEAE: Minnesota Therapeutic Education Alternative 
* Subject: Classifies each test that is given to a student based on the areas measured. There are 8 factors;
  + A: ACCESS Test
  + C: ACCESS Test
  + L: Listening/Speaking
  + M: Math
  + O: ACCESS Test
  + R: Reading
  + S: Science
  + W: Writing
* GradeTested: Grade student was enrolled in when performing assessment.
* AchievementLevel: Used to count the number of students in each achievement level by group such as race, gender, and ethnicity. It must be calculated from other students at a similar graden and subject for a specific year. 
  + ACCESS students are graded on a language proficiency scale of 1 through 6.
  + All-ACCESS students are graded using a P1 through P3 and A1 through A3 scale.
  + There are 4 different levels for the MCA test
    + D: Does not meet standards
    + E: Exceeds standards
    + M: Meets Standards
    + P: Partially meets standards
  + SOLOM tests are 1-5 scale with 5 being the highest
  + TEAE tests are 1-5 scale.     
  