---
title: "CTEDemographics"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(readxl)
library(janitor)
library(lubridate)
library(systemfonts)
reset_font_cache()
library(ggtext)
library(gmodels)
```

```{r themes and join docs, include=FALSE}
theme_line <- theme_bw() +
  theme(legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.text = element_text(margin = margin(l = 2)),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey70", size = 0.1),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))

theme_sf <- theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "white"),
        panel.border = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))

mn_counties <- st_read("Data/Shapefiles/county shapefiles/MNCounties_MNDOT.shp", quiet = TRUE) %>%
  ms_simplify(keep = .01, keep_shapes = TRUE) %>%
  rename(countyfp = FIPS_CODE)

counties.regions <- read_csv("Data/Join docs/county_regions.csv") %>%
  rename(mif = `MIF Region`) %>%
  mutate(countyfp = formatC(countyfp, width = 3, flag = "0"),
         Name = str_to_title(Name),
         Name = str_replace(Name, "Q", "q"),
         Name = str_replace(Name, "Of The", "of the"),
         Name = str_replace(Name, "Mcleod", "McLeod"),
         Dem_Desc = ifelse(Name == "Minnesota", "Minnesota", Dem_Desc) ,
         edr = str_replace(edr, "  ", " "),
         planning.region = str_replace(planning.region, " Minnesota", ""),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"),
         mif = ifelse(is.na(mif), "TC", mif),
         mif = as.factor(mif),
         mif = fct_relevel(mif, "NW", "NE", "WC", "EC", "SW", "SE", "TC"))

```

<br>

# Data prep

Now I'm going to prep the CTE demographics dataset to join with the master.

<br>

```{r cte demographics original}
cte.demographics.original <- read_csv("Data/SLEDS/CTE/CTE_Demographics.csv")  %>%
  mutate(PersonID = as.integer(PersonID)) 

head(cte.demographics.original)

names(cte.demographics.original)

```

<br> 

The original dataset provides whether the student was a basic participant, basic concentrator, basic completor, and/or a centrentrator 150 when taking a CTE course. It looks like each observation is unique by grade of the student. This means that one PersonID can have multiple observations.

In order to determine how I want to organize this data so that each observation represents the full characteristics of each PersonID I need to understand how these values are defined.

* Basic Participant: Any grade 9-12 student who successfully completes one or more state-approved CTE course(s).
* Basic Concentrator: this is the old standard for years and required the student complete 240 hours. Any student successfully enrolled in a single CTE field for more than 240 hours (successful is defined as passing grade). 
* Basic Completor: any concentrator who attains academic and technical knowledge and skills and graduates from high school.
* Concentrator 150: This is the new standard - Any grade 9-12 student who successfully completes 150+ course hours (or more) as part of a state-approved secondary CTE program, within one career field.

Judging by these definitions, it seems that we will want three categories;

1. Those that participated in a CTE course but were not a concentrator or a completor, and
2. Those that participated in a CTE course and graduated as a concentrator.
3. Those that did not participate in a CTE course.

So essentially, filtering this dataset to provide the "highest level achieved" in terms of CTE coursework. We will focus on filtering this dataset to the highest level achieved and will then join with the graduate list which will have all the individuals that did not participate in any CTE courses.

<br>

```{r cte concentrator}
graduate <- read_csv("Data/SLEDS/Masters/Master-4.csv") %>%
  select(PersonID)

cte.concentrator.cat <- cte.demographics.original %>%
  gather(key = "cte.achievement", value = "yes.or.no", 4:7) %>%
  filter(yes.or.no != "Null") %>%
  filter(yes.or.no != "NULL") %>%
  filter(yes.or.no != "N") %>%
  mutate(cte.achievement = as.factor(cte.achievement),
         yes.or.no = as.factor(yes.or.no),
         cte.achievement.code = ifelse(cte.achievement == "BasicParticipant", 0, 1)) %>%
  select(PersonID, cte.achievement, cte.achievement.code) %>%
  group_by(PersonID) %>%
  summarise(cte.achievement.code = sum(cte.achievement.code)) %>%
  ungroup() %>%
  mutate(cte.achievement = ifelse(cte.achievement.code == 0, "CTE Participant", "CTE Concentrator or Completor")) %>%
  select(PersonID, cte.achievement) %>%
  right_join(graduate, by = "PersonID") %>%
  mutate(cte.achievement = ifelse(is.na(cte.achievement), "No CTE", cte.achievement))

head(cte.concentrator.cat)

names(cte.concentrator.cat)
```

<br>

When joined with the master file, there were `r comma(cte.concentrator.cat %>% filter(cte.achievement != "No CTE") %>% nrow(), accuracy = 1)` that were either a CTE participant or concentrator. This is below the ~34,000 that took at least one CTE course according to the CTE enrollment dataset. This could mean that there roughly 5,000 students that took a CTE course but did not graduate.

Lets join this dataset with the master to confirm this thought.

<br>

```{r cte concentrator join master}
master.5 <- read_csv("Data/SLEDS/Masters/Master-4.csv") %>%
  left_join(cte.concentrator.cat, by = "PersonID")

head(master.5)

names(master.5)

```

<br>

After joining with the master dataset, we have `r comma(master.5 %>% nrow(), accuracy = 1)` rows and `r comma(master.5 %>% ncol(), accuracy = 1)` columns.

<br>

# Summary of CTE demographics

We will summarize the CTE demographics data by rural-ness and region to see if there's any significant differences.

First, lets see what percentage of the graduates by cte demographics categories.

The table below shows that of the total graduates from Southwest high schools, nearly half (47%) were a CTE concentrator or completor. This is a bit surprising. Lets see if there are any differences by rural-ness.


<br>

```{r summary cte demographics total}
cte.dem.summary.total <- master.5 %>%
  mutate(cte.achievement = as.factor(cte.achievement)) %>%
  tabyl(cte.achievement)

datatable(cte.dem.summary.total, class = "cell-border stripe", filter = "top", rownames = FALSE) %>%
  formatPercentage(3) %>%
  formatCurrency(2, "", digits = 0)

```

<br>

# {.unnumbered .unlisted .toc-ignore .tabset}

The cross-tabs show a significant difference between the percentage of students by categorization and RUCA categories. Our most rural counties have a significantly higher percentage of graduates that are CTE concentrators/completors (55%) than the other RUCA categories. The percentage decreases as the county group becomes more urban - the urban/town/rural mix county group had 38% of their graduates categorized as a CTE concentrator or completor. 

Our most rural counties also has a smaller percentage of graduates that took no CTE courses (17%) compared to the other RUCA groups. Urban/town/rural mix had the highest percentage with 28%. 

<br>

## RUCA Table

```{r summary cte demographics ruca table}
cte.dem.summary.ruca.table <- master.5 %>%
  mutate(cte.achievement = as.factor(cte.achievement)) %>%
  tabyl(Dem_Desc, cte.achievement) %>%
  adorn_percentages("row")

datatable(cte.dem.summary.ruca.table, class = "cell-border stripe", filter = "top", rownames = FALSE) %>%
  formatPercentage(2:4) 
```

## RUCA Cross-tab

```{r summary cte demographics ruca crosstabs}

CrossTable(master.5$Dem_Desc, master.5$cte.achievement, expected = TRUE, prop.t = FALSE, prop.c = FALSE, prop.chisq = FALSE)


```

<br>

# {.unnumbered .unlisted .toc-ignore .tabset}

Next we will take a look to see if there are any differences by EDR.

Again we see a statistically significant difference in the cross-tabulations. EDR 6W has the most rural counties within it's region so it's not surprising to see that they have the highest percentage of graduates categorized as a CTE completor or concentrator. The lower percentage is EDR 6E which has Mankato.

<br>

## EDR Table

```{r summary cte demographics edr table}
cte.dem.summary.edr.table <- master.5 %>%
  mutate(cte.achievement = as.factor(cte.achievement)) %>%
  tabyl(edr, cte.achievement) %>%
  adorn_percentages("row")

datatable(cte.dem.summary.edr.table, class = "cell-border stripe", filter = "top", rownames = FALSE) %>%
  formatPercentage(2:4) 
```


## EDR cross-tabs

```{r summary cte demographics edr crosstabs}

CrossTable(master.5$edr, master.5$cte.achievement, expected = TRUE, prop.t = FALSE, prop.c = FALSE, prop.chisq = FALSE)

```

# {.unnumbered .unlisted .toc-ignore .tabset}

<br>

Lets take a look at the county map.

The cross tabulations show that there is a statistically significant difference in the percent of graduates in each CTE category by county. 

Swift, Murray, Big Stone and Lac qui Parle counties all have over 50% of their graduates categorized as a CTE concentrator or completor compared to counties such as Pipestone and McLeod who have less than 40%. In fact, Swift and Murray have over 70% of their graduates with the CTE concentrator or completor category. 

<br>

## County map

```{r summary cte demographics county map}
cte.dem.summary.county.map <- master.5 %>%
  mutate(cte.achievement = as.factor(cte.achievement)) %>%
  tabyl(county.name, cte.achievement) %>%
  adorn_percentages("row") %>%
  gather(key = "cte.category", value = "pct", 2:4) %>%
  left_join(counties.regions[,c(1,2)], by = c("county.name" = "Name")) %>%
  left_join(mn_counties[,c(5,7)], by = "countyfp")

cte.dem.summary.county.map.plot<- ggplot(cte.dem.summary.county.map) +
  facet_wrap(~cte.category, ncol = 2) +
  geom_sf_interactive(color = "grey85", aes(geometry = geometry, fill = pct, data_id = countyfp, tooltip = paste(county.name, "\nCTE achievement category: ", cte.category, "\nPercent of graduates: ", percent(pct, accuracy = .1), sep = ""))) +
  theme_sf+
  scale_fill_fermenter(palette = "PuBu", direction = 1, labels = scales::percent) +
  labs(title = "Percent of graduates in each county location of high school by\nCTE achievement\n") +
  theme(legend.box.margin = margin(50, 0, 0, -100),
        text = element_text(size = 18),
        plot.title.position = "plot")

girafe(ggobj = cte.dem.summary.county.map.plot, height_svg = 10, width_svg = 10) %>%
  girafe_options(opts_sizing(rescale = FALSE))


```

## County table

```{r summary cte demographics county table}
cte.dem.summary.county.table <- master.5 %>%
  mutate(cte.achievement = as.factor(cte.achievement)) %>%
  tabyl(county.name, cte.achievement) %>%
  adorn_percentages("row")

datatable(cte.dem.summary.county.table, class = "cell-border stripe", filter = "top", rownames = FALSE) %>%
  formatPercentage(2:4) 
```

## County cross-tabs

```{r summary cte demographics county crosstabs}

CrossTable(master.5$county.name, master.5$cte.achievement, expected = TRUE, prop.t = FALSE, prop.c = FALSE, prop.chisq = FALSE)

```

# {.unnumbered .unlisted .toc-ignore .tabset}

<br>

The next thing to check is to determine whether there is a significant difference by year.

The cross-tabs indicate a relationship, but this is largely due to the significantly large percentage in 2008 and 2009. Both of these years had roughly 70% of the graduates as a CTE concentrator or completor. This is odd and I'll have to show this to Meredith to see if it makes sense.

<br>

## Year table

```{r summary cte demographics year table}
cte.dem.summary.year.table <- master.5 %>%
  mutate(cte.achievement = as.factor(cte.achievement)) %>%
  tabyl(grad.year, cte.achievement) %>%
  adorn_percentages("row")

datatable(cte.dem.summary.year.table, class = "cell-border stripe", filter = "top", rownames = FALSE) %>%
  formatPercentage(2:4) 
```

## Year cross-tab

```{r summary cte demographics year crosstabs}

CrossTable(master.5$grad.year, master.5$cte.achievement, expected = TRUE, prop.t = FALSE, prop.c = FALSE, prop.chisq = FALSE)

```

# {.unnumbered .unlisted .toc-ignore .tabset}

<br>


```{r master 5}
write_csv(master.5, "Data/SLEDS/Masters/Master-5.csv")
```
