---
title: "CTEEnrollment"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(extrafont)
library(readxl)
library(janitor)
library(lubridate)

loadfonts()

```

```{r themes and join docs, include=FALSE}
theme_line <- theme_bw() +
  theme(legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.text = element_text(margin = margin(l = 2)),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey70", size = 0.1),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))

theme_sf <- theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "white"),
        panel.border = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))

mn_counties <- st_read("Data/Shapefiles/county shapefiles/MNCounties_MNDOT.shp", quiet = TRUE) %>%
  ms_simplify(keep = .01, keep_shapes = TRUE) %>%
  rename(countyfp = FIPS_CODE)

counties.regions <- read_csv("Data/Join docs/county_regions.csv") %>%
  rename(mif = `MIF Region`) %>%
  mutate(countyfp = formatC(countyfp, width = 3, flag = "0"),
         Name = str_to_title(Name),
         Name = str_replace(Name, "Q", "q"),
         Name = str_replace(Name, "Of The", "of the"),
         Name = str_replace(Name, "Mcleod", "McLeod"),
         Dem_Desc = ifelse(Name == "Minnesota", "Minnesota", Dem_Desc) ,
         edr = str_replace(edr, "  ", " "),
         planning.region = str_replace(planning.region, " Minnesota", ""),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"),
         mif = ifelse(is.na(mif), "TC", mif),
         mif = as.factor(mif),
         mif = fct_relevel(mif, "NW", "NE", "WC", "EC", "SW", "SE", "TC"))

```

<br>

# Data prep

The next dataset is individuals that were enrolled in CTE courses.

<br>

```{r cte enrollment original}
cte.enrollment.original <- read_csv("Data/SLEDS/CTE/CTE_CourseEnrollments.csv") %>%
  mutate(PersonID = as.integer(PersonID)) %>%
  filter(FiscalYearTimeID != "NULL")

head(cte.enrollment.original)

names(cte.enrollment.original)
```

<br>

This dataset provides each CTE course taken by a student. This means that there were `r comma(nrow(cte.enrollment.original), accuracy = 1)` instances of a student taking a CTE course. There are `r comma(ncol(cte.enrollment.original), accuracy = 1)` columns providing the PersonID taking the course, the fiscal year, the K12 organization, along with details about the actual course itself. 

There are two goals with this dataset. 

1. Format it so each observation represents a single PersonID along with the total number of CTE courses they took.
2. Format it so each observation represents a single PersonID along with the number of CTE courses taken in each career field.

I could use career cluster but there are 20 different career clusters. The career fields are more broad and there are only 6 categories.

<br>

```{r cteenrollment n total cte courses}
graduate <- read_csv("Data/SLEDS/Masters/Master-3.csv") %>%
  select(PersonID)

cte.n.total.courses <- cte.enrollment.original %>%
  group_by(PersonID) %>%
  summarise(total.cte.courses.taken = n()) %>%
  ungroup() %>%
  right_join(graduate, by = "PersonID") %>%
  mutate(total.cte.courses.taken = ifelse(is.na(total.cte.courses.taken), 0, total.cte.courses.taken))

head(cte.n.total.courses)

names(cte.n.total.courses)
```

<br>

A total of `r cte.n.total.courses %>% filter(total.cte.courses.taken > 0) %>% nrow() %>% scales::comma(accuracy = 1) ` individuals took at least one CTE course in the dataset. 

Next I'm going to create a dataset that calculates the number of CTE courses each individual has taken by CareerField. 

<br>

```{r cteenrollment by careerfield}
cte.n.careerfield <- cte.enrollment.original %>%
  mutate(CareerField = str_replace(CareerField, "Information Technology", "6")) %>%
  group_by(PersonID, CareerField) %>%
  summarise(cte.n.careerfield = n()) %>%
  ungroup() %>%
  right_join(graduate, by = "PersonID") %>%
  complete(PersonID, CareerField, fill = list(cte.n.careerfield = 0)) %>%
  drop_na(CareerField) %>%
  mutate(CareerField = paste("cte.careerfield.", CareerField, sep = "")) %>%
  spread(key = CareerField, value = cte.n.careerfield)

head(cte.n.careerfield)

names(cte.n.careerfield)
```

<br>

There seems to be a minor coding error in the careerfield column. For one particular class, it was consistently labeled "Information Technology" even though the field requires a number value. In addition, "information Technology" doesn't seem to match the course due to the following information;

Program Course Name: Family and Consumer Science
Course title: Apparrel
CareerCluster: Design
CareerFieldName: 6 (which should be a letter value, not a code).

Due to this information, I recoded anything that was "Information Technology" for that particular course as a "6".

After doing this, the dataset has all individuals/graduates with a column for each CTE course careerfield containing the number of courses they took within that CareerField. I will summarise below to see how many students took courses within each CareerField.

Now lets join the CTE enrollment data with the master dataset.

<br>

```{r join cte enrollment with master}
master.3 <- read_csv("Data/SLEDS/Masters/Master-3.csv")

master.4 <- master.3 %>%
  left_join(cte.n.total.courses, by = "PersonID") %>%
  left_join(cte.n.careerfield, by = "PersonID")

head(master.4)

names(master.4)
```

<br>

After joining the CTE enrollment data with the master dataset we continue to have `r master.4 %>% nrow() %>% comma(accuracy = 1)` with an updated number of `r master.4 %>% ncol() %>% comma(accuracy = 1)` columns.

<br>

# Summary of CTE enrollment data

<br>

## Summary - Number of CTE courses taken by a student

<br>

We'll start with summarising the number of cte courses the observations have taken.

<br>

```{r summarise n CTE courses taken, echo=FALSE}
cte.n.courses.summary.total <- master.4 %>%
  summarise(mean = mean(total.cte.courses.taken),
            median = median(total.cte.courses.taken),
            sd = sd(total.cte.courses.taken),
            n = comma(n(), accuracy = 1),
            min = min(total.cte.courses.taken),
            max = max(total.cte.courses.taken)) %>%
  mutate(population = "All graduates",
         mean = comma(mean, accuracy = .01),
         sd = comma(sd, accuracy = .01))

cte.n.courses.summary.ctestudents <- master.4 %>%
  filter(total.cte.courses.taken > 0) %>%
  summarise(mean = mean(total.cte.courses.taken),
            median = median(total.cte.courses.taken),
            sd = sd(total.cte.courses.taken),
            n = comma(n(), accuracy = 1),
            min = min(total.cte.courses.taken),
            max = max(total.cte.courses.taken)) %>%
  mutate(population = "Only graduates with CTE courses > 0",
         mean = comma(mean, accuracy = .01),
         sd = comma(sd, accuracy = .01))
  
cte.n.courses.summary.total.ctestudents <- cte.n.courses.summary.total %>%
  rbind(cte.n.courses.summary.ctestudents) %>%
  select(population, n, mean, median, sd, min, max)

datatable(cte.n.courses.summary.total.ctestudents, class = "cell-border stripe", filter = "top")

```

<br>

The average number of CTE courses taken by all PersonID in the master dataset is 4.98 with a standard deviation of +/-4.9. This is a high standard deviation meaning there are values quite far from the mean. 

The average number of CTE courses taken by all PersonID that have taken at least one CTE course was 6.10 with a standard deviation of +/-4.75. So the standard deviation decreases a bit with this population likely because we got rid of all the "0" values that are in the full dataset.

The highest number of CTE courses taken is 69. That is quite high so I will check with Meredith on that one.

Lets look at the distribution to see if that value is an outlier.



<br>

```{r summarise n CTE courses taken distribution, echo=FALSE}
master.4.hist <- master.4 %>%
  mutate(key = seq(n())) %>%
  select(key, total.cte.courses.taken)

cte.n.courses.taken.dist.plot <- ggplot(master.4.hist, aes(total.cte.courses.taken)) +
  geom_histogram_interactive(aes(data_id = key, tooltip = paste("N = ", comma(..count.., accuracy = 1), "\nPercent: ", percent(..count.. / sum(..count..), accuracy = .1)))) +
  labs(x="", y = "", color="", title = "Histogram - N CTE courses")+
  scale_y_continuous(labels=scales::comma)+
  theme_line+
  theme(legend.position = "bottom",
        text = element_text(size = 18))

  girafe(ggobj = cte.n.courses.taken.dist.plot, width_svg = 10, height_svg = 7) %>%
    girafe_options(opts_selection(type = "none"),
                   opts_toolbar(saveaspng = FALSE),
                   opts_sizing(rescale = FALSE))


```

<br>

Up next is to see if the percentage of students taking a CTE course varies significantly by RUCA group.

<br>

```{r cte n courses ruca, echo=FALSE}
cte.yes.ruca <- master.4 %>%
  mutate(cte.yes = ifelse(total.cte.courses.taken > 0, "Yes", "No")) %>%
  tabyl(Dem_Desc, cte.yes) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 1)

cte.n.ruca <- master.4 %>%
  filter(total.cte.courses.taken > 0 ) %>%
  group_by(Dem_Desc) %>%
  summarise(mean = comma(mean(total.cte.courses.taken), accuracy = .1),
            median = comma(median(total.cte.courses.taken), accuracy = 1),
            sd = comma(sd(total.cte.courses.taken), accuracy = .1),
            n = comma(n(), accuracy = 1),
            min = comma(min(total.cte.courses.taken), accuracy = 1),
            max = comma(max(total.cte.courses.taken), accuracy = 1)) %>%
  ungroup()

datatable(cte.yes.ruca, class = "cell-border stripe", filter = "top")

tags$br()

datatable(cte.n.ruca, class = "cell-border stripe", filter = "top")

```

<br>

There isn't significant variation by RUCA group. About 80% of the students within in each RUCA group have taken at least 1 CTE course. The percentage does increase as a county is categorized more rural.

Of those students that took at least 1 CTE course, the average number of CTE courses taken ranged from 5.2 in urban/town/rural mix counties to 6.3 in town/rural mix county group. 

Now lets check everything by planning region.

<br>

```{r cte n courses planning region, echo=FALSE}
cte.yes.edr <- master.4 %>%
  mutate(cte.yes = ifelse(total.cte.courses.taken > 0, "Yes", "No")) %>%
  tabyl(edr, cte.yes) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 1)

cte.n.edr <- master.4 %>%
  filter(total.cte.courses.taken > 0 ) %>%
  group_by(edr) %>%
  summarise(mean = comma(mean(total.cte.courses.taken), accuracy = .1),
            median = comma(median(total.cte.courses.taken), accuracy = 1),
            sd = comma(sd(total.cte.courses.taken), accuracy = .1),
            n = comma(n(), accuracy = 1),
            min = comma(min(total.cte.courses.taken), accuracy = 1),
            max = comma(max(total.cte.courses.taken), accuracy = 1)) %>%
  ungroup()

datatable(cte.yes.edr, class = "cell-border stripe", filter = "top")

tags$br()

datatable(cte.n.ruca, class = "cell-border stripe", filter = "top")

```

<br>

## Summary - Career field of CTE courses taken

<br>

Now we will take a look at the breakdown of CTE courses taken by career field for each PersonID. I expect there to be a lot of variation in this data since different schools might focus on different types of CTE offerings depending on the local employment makeup.

First, lets look at the concentrations of courses taken across career fields.

<br>

```{r cte career fields concentration, echo=FALSE}
cte.cf.concentration.total <- master.4 %>%
  select(PersonID, cte.careerfield.1, cte.careerfield.2, cte.careerfield.3, cte.careerfield.4, cte.careerfield.5, cte.careerfield.6, cte.careerfield.7, cte.careerfield.8, cte.careerfield.9, cte.careerfield.NULL) %>%
  gather(key = "cte.careerfield", value = "n.courses", 2:11) %>%
  group_by(cte.careerfield) %>%
  summarise(n = sum(n.courses)) %>%
  ungroup() %>%
  mutate(pct = percent(n / sum(n), accuracy = .1))

names(cte.cf.concentration.total)
```


