---
title: "CTEEnrollment"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(readxl)
library(janitor)
library(lubridate)
library(systemfonts)
reset_font_cache()
library(ggtext)
library(gmodels)


```

```{r themes and join docs, include=FALSE}
theme_line <- theme_bw() +
  theme(legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.text = element_text(margin = margin(l = 2)),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey70", size = 0.1),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))

theme_sf <- theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "white"),
        panel.border = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))

mn_counties <- st_read("Data/Shapefiles/county shapefiles/MNCounties_MNDOT.shp", quiet = TRUE) %>%
  ms_simplify(keep = .01, keep_shapes = TRUE) %>%
  rename(countyfp = FIPS_CODE)

counties.regions <- read_csv("Data/Join docs/county_regions.csv") %>%
  rename(mif = `MIF Region`) %>%
  mutate(countyfp = formatC(countyfp, width = 3, flag = "0"),
         Name = str_to_title(Name),
         Name = str_replace(Name, "Q", "q"),
         Name = str_replace(Name, "Of The", "of the"),
         Name = str_replace(Name, "Mcleod", "McLeod"),
         Dem_Desc = ifelse(Name == "Minnesota", "Minnesota", Dem_Desc) ,
         edr = str_replace(edr, "  ", " "),
         planning.region = str_replace(planning.region, " Minnesota", ""),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"),
         mif = ifelse(is.na(mif), "TC", mif),
         mif = as.factor(mif),
         mif = fct_relevel(mif, "NW", "NE", "WC", "EC", "SW", "SE", "TC"))

```

```{r cte careerfields joined with names}
cte.careerfield.names <- read_csv("Data/SLEDS/CTE/career-field-names.csv") %>%
  select(1,2) %>%
  mutate(cf.code = paste("cte.careerfield.", cf.code, sep = ""))
  

```

<br>

# Data prep

The next dataset is individuals that were enrolled in CTE courses.

<br>

```{r cte enrollment original}
cte.enrollment.original <- read_csv("Data/SLEDS/CTE/CTE_CourseEnrollments.csv") %>%
  mutate(PersonID = as.integer(PersonID)) %>%
  filter(FiscalYearTimeID != "NULL")

head(cte.enrollment.original)

names(cte.enrollment.original)
```

<br>

This dataset provides each CTE course taken by a student. This means that there were `r comma(nrow(cte.enrollment.original), accuracy = 1)` instances of a student taking a CTE course. There are `r comma(ncol(cte.enrollment.original), accuracy = 1)` columns providing the PersonID taking the course, the fiscal year, the K12 organization, along with details about the actual course itself.

There are two goals with this dataset.

1.  Format it so each observation represents a single PersonID along with the total number of CTE courses they took.
2.  Format it so each observation represents a single PersonID along with the number of CTE courses taken in each career field.

I could use career cluster but there are 20 different career clusters. The career fields are more broad and there are only 6 categories.

<br>

```{r cteenrollment n total cte courses}
graduate <- read_csv("Data/SLEDS/Masters/Master-3.csv") %>%
  select(PersonID)

cte.n.total.courses <- cte.enrollment.original %>%
  group_by(PersonID) %>%
  summarise(total.cte.courses.taken = n()) %>%
  ungroup() %>%
  right_join(graduate, by = "PersonID") %>%
  mutate(total.cte.courses.taken = ifelse(is.na(total.cte.courses.taken), 0, total.cte.courses.taken))

head(cte.n.total.courses)

names(cte.n.total.courses)
```

<br>

A total of `r cte.n.total.courses %>% filter(total.cte.courses.taken > 0) %>% nrow() %>% scales::comma(accuracy = 1)` individuals took at least one CTE course in the dataset.

Next I'm going to create a dataset that calculates the number of CTE courses each individual has taken by CareerField.

<br>

```{r cteenrollment by careerfield}
cte.n.careerfield <- cte.enrollment.original %>%
  mutate(CareerField = str_replace(CareerField, "Information Technology", "6")) %>%
  group_by(PersonID, CareerField) %>%
  summarise(cte.n.careerfield = n()) %>%
  ungroup() %>%
  right_join(graduate, by = "PersonID") %>%
  complete(PersonID, CareerField, fill = list(cte.n.careerfield = 0)) %>%
  drop_na(CareerField) %>%
  mutate(CareerField = paste("cte.careerfield.", CareerField, sep = "")) %>%
  spread(key = CareerField, value = cte.n.careerfield)

head(cte.n.careerfield)

names(cte.n.careerfield)
```

<br>

There seems to be a minor coding error in the careerfield column. For one particular class, it was consistently labeled "Information Technology" even though the field requires a number value. In addition, "information Technology" doesn't seem to match the course due to the following information;

Program Course Name: Family and Consumer Science Course title: Apparrel CareerCluster: Design CareerFieldName: 6 (which should be a letter value, not a code).

Due to this information, I recoded anything that was "Information Technology" for that particular course as a "6".

After doing this, the dataset has all individuals/graduates with a column for each CTE course careerfield containing the number of courses they took within that CareerField. I will summarise below to see how many students took courses within each CareerField.

Now lets join the CTE enrollment data with the master dataset.

<br>

```{r join cte enrollment with master}
master.3 <- read_csv("Data/SLEDS/Masters/Master-3.csv")

master.4 <- master.3 %>%
  left_join(cte.n.total.courses, by = "PersonID") %>%
  left_join(cte.n.careerfield, by = "PersonID")

head(master.4)

names(master.4)
```

<br>

After joining the CTE enrollment data with the master dataset we continue to have `r master.4 %>% nrow() %>% comma(accuracy = 1)` with an updated number of `r master.4 %>% ncol() %>% comma(accuracy = 1)` columns.

<br>

# Summary of CTE enrollment data

<br>

## Summary - Number of CTE courses taken by a student

<br>

We'll start with summarising the number of cte courses the observations have taken.

The tables below show that the average number of CTE courses taken by all PersonID in the master dataset is 5.69 with a standard deviation of +/-4.95. This is a high standard deviation meaning there are values quite far from the mean.

The second table below shows the average number of CTE courses taken by all PersonID that have taken at least one CTE course was 6.29 with a standard deviation of +/-4.82. So the standard deviation decreases a bit with this population likely because we got rid of all the "0" values that are in the full dataset. But not significant.

<br>

```{r summarise n CTE courses taken, echo=FALSE}
cte.n.courses.summary.total <- master.4 %>%
  summarise(mean = mean(total.cte.courses.taken),
            median = median(total.cte.courses.taken),
            sd = sd(total.cte.courses.taken),
            n = comma(n(), accuracy = 1),
            min = min(total.cte.courses.taken),
            max = max(total.cte.courses.taken)) %>%
  mutate(population = "All graduates",
         mean = comma(mean, accuracy = .01),
         sd = comma(sd, accuracy = .01))

cte.n.courses.summary.ctestudents <- master.4 %>%
  filter(total.cte.courses.taken > 0) %>%
  summarise(mean = mean(total.cte.courses.taken),
            median = median(total.cte.courses.taken),
            sd = sd(total.cte.courses.taken),
            n = comma(n(), accuracy = 1),
            min = min(total.cte.courses.taken),
            max = max(total.cte.courses.taken)) %>%
  mutate(population = "Only graduates with CTE courses > 0",
         mean = comma(mean, accuracy = .01),
         sd = comma(sd, accuracy = .01))
  
cte.n.courses.summary.total.ctestudents <- cte.n.courses.summary.total %>%
  rbind(cte.n.courses.summary.ctestudents) %>%
  select(population, n, mean, median, sd, min, max)

datatable(cte.n.courses.summary.total.ctestudents, class = "cell-border stripe", filter = "top")

```

<br>

Below is a chart showing the distribution of CTE courses taken by individuals. This shows that a a large majority of students take 10 courses or less. 

<br>

```{r summarise n CTE courses taken distribution, echo=FALSE}
master.4.hist <- master.4 %>%
  mutate(key = seq(n())) %>%
  select(key, total.cte.courses.taken)

cte.n.courses.taken.dist.plot <- ggplot(master.4.hist, aes(total.cte.courses.taken)) +
  geom_histogram_interactive(aes(data_id = key, tooltip = paste("N = ", comma(..count.., accuracy = 1), "\nPercent: ", percent(..count.. / sum(..count..), accuracy = .1)))) +
  labs(x="", y = "", color="", title = "Histogram - N CTE courses")+
  scale_y_continuous(labels=scales::comma)+
  theme_line+
  theme(legend.position = "bottom",
        text = element_text(size = 18))

  girafe(ggobj = cte.n.courses.taken.dist.plot, width_svg = 10, height_svg = 7) %>%
    girafe_options(opts_selection(type = "none"),
                   opts_toolbar(saveaspng = FALSE),
                   opts_sizing(rescale = FALSE))


```
## {.unnumbered .unlisted .toc-ignore .tabset}

<br>

Up next is to see if the percentage of students taking a CTE course varies significantly by RUCA group.

There is a significant difference in whether a student has taken at least one CTE course and their high school county RUCA category.  significant variation by RUCA group. Although 90.3% of all graduates took at least one CTE course, that percentage is highest in entirely rural counties and decreases as counties become more urban. 

Of those students that took at least 1 CTE course, the average number of CTE courses taken ranged from 5.4 in urban/town/rural mix counties to 6.5 in town/rural mix county group.

<br>

### RUCA table

```{r taken at least one cte course table ruca, echo=FALSE}
cte.yes.ruca <- master.4 %>%
  mutate(cte.yes = ifelse(total.cte.courses.taken > 0, "Yes", "No")) %>%
  tabyl(Dem_Desc, cte.yes) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 1)

cte.n.ruca <- master.4 %>%
  filter(total.cte.courses.taken > 0 ) %>%
  group_by(Dem_Desc) %>%
  summarise(mean = comma(mean(total.cte.courses.taken), accuracy = .1),
            median = comma(median(total.cte.courses.taken), accuracy = 1),
            sd = comma(sd(total.cte.courses.taken), accuracy = .1),
            n = comma(n(), accuracy = 1),
            min = comma(min(total.cte.courses.taken), accuracy = 1),
            max = comma(max(total.cte.courses.taken), accuracy = 1)) %>%
  ungroup()

datatable(cte.yes.ruca, class = "cell-border stripe", filter = "top")

tags$br()

datatable(cte.n.ruca, class = "cell-border stripe", filter = "top")

```

### RUCA cross-tabs

```{r taken at least 1 cte course ruca cross tabs}
cte.yes.no.ruca <- master.4 %>%
  mutate(cte.yes = ifelse(total.cte.courses.taken > 0, "Yes", "No")) 



CrossTable(cte.yes.no.ruca$Dem_Desc, cte.yes.no.ruca$cte.yes, expected = TRUE, prop.t = FALSE, prop.c = FALSE, prop.chisq = FALSE)
```

## {.unnumbered .unlisted .toc-ignore .tabset}


<br>

Now lets check everything by EDR.

Not surprisingly, there is a significant difference in the percentage of students who took at least one CTE course and their EDR.

The more rural EDR (EDR 6W) had the highest proportion of graduates who have taken at least one CTE course with 93%. The lowest was in the EDR with more urban areas - EDR 8 - which had 89%.

Of the students that have taken at least one CTE course, the average number was lowest in EDR 8 with 6.0 and highest in EDR 6W with 7.2. 

<br>

### EDR Tables
```{r cte n courses planning region, echo=FALSE}
cte.yes.edr <- master.4 %>%
  mutate(cte.yes = ifelse(total.cte.courses.taken > 0, "Yes", "No")) %>%
  tabyl(edr, cte.yes) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 1)

cte.n.edr <- master.4 %>%
  filter(total.cte.courses.taken > 0 ) %>%
  group_by(edr) %>%
  summarise(mean = comma(mean(total.cte.courses.taken), accuracy = .1),
            median = comma(median(total.cte.courses.taken), accuracy = 1),
            sd = comma(sd(total.cte.courses.taken), accuracy = .1),
            n = comma(n(), accuracy = 1),
            min = comma(min(total.cte.courses.taken), accuracy = 1),
            max = comma(max(total.cte.courses.taken), accuracy = 1)) %>%
  ungroup()

datatable(cte.yes.edr, class = "cell-border stripe", filter = "top")

tags$br()

datatable(cte.n.edr, class = "cell-border stripe", filter = "top")

```

### EDR Cross-tabs

```{r taken at least 1 cte course edr cross tabs}
cte.yes.no.edr <- master.4 %>%
  mutate(cte.yes = ifelse(total.cte.courses.taken > 0, "Yes", "No")) 



CrossTable(cte.yes.no.edr$edr, cte.yes.no.edr$cte.yes, expected = TRUE, prop.t = FALSE, prop.c = FALSE, prop.chisq = FALSE)
```

## {.unnumbered .unlisted .toc-ignore}


## Summary - Career field of CTE courses taken

Let's dig into the breadth of CTE course concentrations a graduate would take by looking at the proportion of graduates that took at least one course within a concentration.

The table below shows that over half of the graduates took at least one course in business, management, and administration; ag, food and natural resources; and human services. 

From there, the percentage drops off quite a bit with engineering, manufacturing, and technology next with 19% and arts, communications and information systems with 7%. Health Science Technology is 1%.


<br>

```{r cte concentration grad prop}
cte.cf.prop <- master.4 %>%
  gather(key = "cf", value = "n.courses", 19:28) %>%
  mutate(cte.cf.taken = ifelse(n.courses > 0, "Yes", "No")) %>%
  tabyl(cf, cte.cf.taken) %>%
  mutate(pct.taken.cf = Yes / (Yes + No)) %>%
  left_join(cte.careerfield.names, by = c("cf" = "cf.code")) %>%
  select(5,2,3,4) %>%
  arrange(desc(pct.taken.cf))

datatable(cte.cf.prop, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(scrollX = TRUE)) %>%
  formatCurrency(2:3, "", digits = 0) %>%
  formatPercentage(4)

```

<br>

Now we will take a look at the breakdown of the number of times a CTE courses has been taken within a career field for each PersonID. I expect there to be a lot of variation in this data since different schools might focus on different types of CTE offerings depending on the local employment makeup.

First, lets look at the concentrations of courses taken across career fields.

Not surprisingly, Agriculture, Food & Natural Resources was the careerfield with the highest percentage of courses taken with 34.9%. This was followed by business, management & administration with 24.9% and human services with 20.6% of CTE courses taken in that career field.

<br>


```{r cte career fields concentration, echo=FALSE}
cte.cf.concentration.total <- master.4 %>%
  select(PersonID, cte.careerfield.1, cte.careerfield.2, cte.careerfield.3, cte.careerfield.4, cte.careerfield.5, cte.careerfield.6, cte.careerfield.7, cte.careerfield.8, cte.careerfield.9, cte.careerfield.NULL) %>%
  gather(key = "cte.careerfield", value = "n.courses", 2:11) %>%
  group_by(cte.careerfield) %>%
  summarise(n = sum(n.courses)) %>%
  ungroup() %>%
  mutate(pct = percent(n / sum(n), accuracy = .1)) %>%
  left_join(cte.careerfield.names, by = c("cte.careerfield" = "cf.code")) %>%
  select(4, 1, 2, 3)

datatable(cte.cf.concentration.total, class = "cell-border stripe", filter = "top", rownames = FALSE)

```

<br>

Next we will see if there's any significant differences by region.

There is one significant difference when looking at the breakdown of CTE courses taken within careerfield by EDR. EDR 8 - Southwest had a significantly higher percentage of courses taken within the Agriculture, Food & Natural Resources career field with 44.1% compared to 28.7% and 26.4% in EDR 6E and 6W respectively.

There was also a significantly lower percentage of courses taken in Engineering, Manufacturing and Technology with 4.0% of courses taken compared to 13.7% in both EDR 6E and 6W.

<br>

```{r cte career fields concentration  by EDR, echo=FALSE}
cte.cf.concentration.edr <- master.4 %>%
  select(PersonID, edr,cte.careerfield.1, cte.careerfield.2, cte.careerfield.3, cte.careerfield.4, cte.careerfield.5, cte.careerfield.6, cte.careerfield.7, cte.careerfield.8, cte.careerfield.9, cte.careerfield.NULL) %>%
  gather(key = "cte.careerfield", value = "n.courses", 3:12) %>%
  group_by(edr, cte.careerfield) %>%
  summarise(n = sum(n.courses)) %>%
  ungroup() %>%
  group_by(edr) %>%
  mutate(pct = percent(n / sum(n), accuracy = .1)) %>%
  left_join(cte.careerfield.names, by = c("cte.careerfield" = "cf.code")) %>%
  select(1,5,4) %>%
  spread(key = edr, value = pct)

  datatable(cte.cf.concentration.edr, class = "cell-border stripe", filter = "top", rownames = FALSE)

```

Lets see how this looks broken down by county.

The county maps do show some differences in percentages. A significantly higher percentage of CTE courses taken were in Jackson (94.6%), Pipestone (65.9%) Cottonwood (56.5%), and Lincoln (53.8%) counties.

There were also higher percentages of courses taken in business, management and administration in Big Stone (54.4%), Murray (43.4%), and Renville (40.4%) counties.

<br>

```{r cte career fields concentration  by county, echo=FALSE}
cte.cf.concentration.county <- master.4 %>%
  select(PersonID, county.name, countyfp, cte.careerfield.1, cte.careerfield.2, cte.careerfield.3, cte.careerfield.4, cte.careerfield.5, cte.careerfield.6, cte.careerfield.7, cte.careerfield.8, cte.careerfield.9, cte.careerfield.NULL) %>%
  gather(key = "cte.careerfield", value = "n.courses", 4:13) %>%
  group_by(county.name, countyfp, cte.careerfield) %>%
  summarise(n = sum(n.courses)) %>%
  ungroup() %>%
  group_by(county.name, countyfp) %>%
  mutate(pct = n / sum(n)) %>%
  left_join(cte.careerfield.names, by = c("cte.careerfield" = "cf.code")) %>%
  select(1,2,6,5) %>%
  left_join(mn_counties[,c(4,7)], by = "countyfp")

cte.cf.concentration.county.map <- ggplot(cte.cf.concentration.county) +
  facet_wrap(~cf.name, ncol = 2,
             labeller = label_wrap_gen(15)) +
  geom_sf_interactive(color = "grey85", aes(geometry = geometry, fill = pct, data_id = countyfp, tooltip = paste(county.name, "\nCareer field: ", cf.name, "\nPercent of courses taken in this careerfield: ", percent(pct, accuracy = .1), sep = ""))) +
  theme_sf+
  scale_fill_fermenter(palette = "PuBu", direction = 1, labels = scales::percent) +
  labs(title = str_wrap("Percent of CTE courses taken within each career field by county", 39)) +
  theme(legend.box.margin = margin(0, 0, 0, 0),
        text = element_text(size = 18))

girafe(ggobj = cte.cf.concentration.county.map, height_svg = 20) %>%
  girafe_options(opts_sizing(rescale = FALSE))

```

<br>

The next big question is whether these percentages vary by year. Are there any trends?

This is honestly a bit surprising. The chart raises a couple of questions.

There was a significant decrease in the NA categorized courses taken. With this was a large increase in the percentage of students taking Ag, food and natural resouces courses. The remaining courses pretty much remained the same.

<br>

```{r cte career fields concentration  by year, echo=FALSE}
cte.cf.concentration.year <- master.4 %>%
  select(PersonID, grad.year, cte.careerfield.1, cte.careerfield.2, cte.careerfield.3, cte.careerfield.4, cte.careerfield.5, cte.careerfield.6, cte.careerfield.7, cte.careerfield.8, cte.careerfield.9, cte.careerfield.NULL) %>%
  gather(key = "cte.careerfield", value = "n.courses", 3:12) %>%
  group_by(grad.year, cte.careerfield) %>%
  summarise(n = sum(n.courses)) %>%
  ungroup() %>%
  group_by(grad.year) %>%
  mutate(pct = n / sum(n)) %>%
  left_join(cte.careerfield.names, by = c("cte.careerfield" = "cf.code")) %>%
  select(1,5,4) %>%
  filter(grad.year > 2006) %>%
  mutate(data_id = seq(n()))

cte.cf.concentration.year.plot <- ggplot(cte.cf.concentration.year, aes(grad.year, pct, color = cf.name)) +
  geom_point_interactive(size = 3, aes(data_id = data_id, tooltip = paste(cf.name, "\nYear: ", grad.year, "\nCareer field: ", cf.name, "\nPercent of CTE courses taken: ", percent(pct, accuracy = .1), sep = ""))) +
  geom_line() +
  labs(x="", y = "", color="", title = "Percent of taken CTE courses")+
  geom_label(data = filter(cte.cf.concentration.year, grad.year == max(cte.cf.concentration.year$grad.year)), aes(label = percent(pct, accuracy = .1)), show.legend = FALSE, size = 5) +
  scale_y_continuous(labels=scales::percent)+
  scale_x_continuous(breaks = seq(1900, 2050, 2),
                     guide = guide_axis(n.dodge = 2)) +
  theme_line+
  scale_color_manual(values= brewer.pal(n = 9, "Set1"),
                     guide = guide_legend(ncol = 3)) +
  theme(legend.position = "bottom",
        text = element_text(size = 18))

  girafe(ggobj = cte.cf.concentration.year.plot, width_svg = 10, height_svg = 7) %>%
    girafe_options(opts_selection(type = "none"),
                   opts_toolbar(saveaspng = FALSE),
                   opts_sizing(rescale = FALSE))


```

<br>


```{r write master}
write_csv(master.4, "Data/SLEDS/Masters/Master-4.csv")
```
