---
title: "CTEEnrollment"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(extrafont)
library(readxl)
library(janitor)
library(lubridate)

loadfonts()

```

<br>

# Data prep

The next dataset is individuals that were enrolled in CTE courses.

<br>

```{r cte enrollment original}
cte.enrollment.original <- read_csv("Data/SLEDS/CTE/CTE_CourseEnrollments.csv") %>%
  mutate(PersonID = as.integer(PersonID)) %>%
  filter(FiscalYearTimeID != "NULL")

head(cte.enrollment.original)

names(cte.enrollment.original)
```

<br>

This dataset provides each CTE course taken by a student. This means that there were `r comma(nrow(cte.enrollment.original), accuracy = 1)` instances of a student taking a CTE course. There are `r comma(ncol(cte.enrollment.original), accuracy = 1)` columns providing the PersonID taking the course, the fiscal year, the K12 organization, along with details about the actual course itself. 

There are two goals with this dataset. 

1. Format it so each observation represents a single PersonID along with the total number of CTE courses they took.
2. Format it so each observation represents a single PersonID along with the number of CTE courses taken in each career field.

I could use career cluster but there are 20 different career clusters. The career fields are more broad and there are only 6 categories.

