---
title: "K12Enrollment"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(readxl)
library(janitor)
library(lubridate)
library(systemfonts)
reset_font_cache()
library(ggtext)


```

<br>

# Data prep

This dataset is of all the individuals that have graduated from a high school in Southwest Minnesota.

<br>

```{r import original dataset}

enrollment.original <- read_csv("Data/SLEDS/K12 Enrollment/K12Enrollment.csv")  %>%
  mutate(PersonID = as.integer(PersonID),
         K12OrganizationID = as.integer(K12OrganizationID)) %>%
  drop_na(PersonID)

head(enrollment.original)

names(enrollment.original)

```


<br>

In the dataset there are `r comma(nrow(enrollment.original), accuracy = 1)` observations along with `r comma(ncol(enrollment.original), accuracy = 1)` columns. Each observation does not represent a unique, unduplicated individual, but rather a unique status for the individual. For example, one individual can have multiple observations if they changed schools, for each school year, or if their economic status changed in the middle of a school year. 

My goal is to end up with a dataset where each observation (row) represents a unique (unduplicated) individual that's tied to the school from which they graduated. Here's how I will do that.

* PersonID: change to integer
* K12OrganizationID: change to integer.
* K12EnrTimeID: Change so it's only the year
* Grade: eliminate - unimportant since we are mostly concerned about whether they graduated.
* EconomicStatusIndicator: I will change this to two categories. Whether they were ever eligible for reduce priced or free meal will be coded as 1 (eligible = codes 1, 2, 4, 5) at any point in the dataset. The remaining will be coded as 0.
* FederalRaceIndicator: eliminate - I'm going to use the RaceEthnicity column instead. 
* PSEO columns: I'm most concerned about whether they participated in PSEO or not. A new column will be created indicated whether there was a "Y" coded for either the "PSEOConcurrentEnrollmentIndicator" or "PostSecondaryEnrollmentOptionsIndicator" columns at any point in the dataset. The two original columns will be eliminated.
* StatusEnd: I will only include individuals that have StatusEnd coded as an 8 or 9.
* Gender: included and not changed.
* SpecialEdStatus: I will split this into two categories. The first is that a student did not require special education services (codes 1-2) at any point in the dataset. The other category will be that they require special education (codes 4 - 9) at any point in the dataset. 
* Limited English: keep with no changes.
* Primary language: I'm going to create two categories. One being that English is the primary language while the other will be that English is NOT the primary language at any point in the dataset.
* Homeless: Keep with no changes.

We first need to create a list of all individuals with the StatusEnd code of 8 or 9 (graduated). In addition, we need to remove any student that graduated before 2008. This is because we will not have their full Sophmore - Senior records which is especially important when attempting to identify whether the number of CTE courses taken predicts migration. 

This "graduates list" will be the list on which we will bind all the other variables.

<br>

```{r k12enrollment StatusEnd}
graduates <- enrollment.original %>%
  filter(StatusEnd %in% c(8,9)) %>%
  mutate(grad.year = str_sub(K12EnrTimeID, 1, 4),
         grad.year = as.integer(grad.year)) %>%
  select(PersonID, K12OrganizationID, grad.year, Gender, LimitedEnglishProficiencyIndicator, HomelessIndicator) %>%
  distinct(PersonID, .keep_all = TRUE) %>%
  filter(grad.year > 2007)


```

<br>

Unfortunately, we still end up with duplicates so I just did a distinct on PersonID. We will send this to Meredith to figure out. 

Next we need to create new columns for the variables that will indicate a certain status has been met at any point in the dataset for each PersonID. For example, we want a column that indicates a PersonID was enrolled in PSEO at some point in the dataset. The variables that need a new column created are;

* EconomicStatusIndicator
* PostSecondaryEnrollmentOptionsIndicator & PSEOConcurrentEnrollmentIndicator
* SpecialEdStatus
* HomeprimaryLanguage

<br>

```{r K12Enrollment creating new columns}

# Lets start with economic status indicator
k12.enr.esi.indicator <- enrollment.original %>%
  select(PersonID, EconomicStatusIndicator) %>%
  group_by(PersonID) %>%
  mutate(economic.status.code = ifelse(EconomicStatusIndicator %in% c(1,2,4,5), 1, 0),
         economic.status.code.sum = sum(economic.status.code)) %>%
  ungroup() %>%
  distinct(PersonID, .keep_all = TRUE) %>%
  mutate(economic.status = ifelse(economic.status.code.sum > 0, 1, 0)) %>%
  select(PersonID, economic.status)

# PSEO

k12.enr.pseo.indicator <- enrollment.original %>%
  select(PersonID, PostSecondaryEnrollmentOptionsIndicator, PSEOConcurrentEnrollmentIndicator) %>%
  mutate(PostSecondaryEnrollmentOptionsIndicator = ifelse(PostSecondaryEnrollmentOptionsIndicator == "Y", 1, 0),
         PSEOConcurrentEnrollmentIndicator = ifelse(PSEOConcurrentEnrollmentIndicator == "Y", 1, 0)) %>%
  gather(key = "pseo.type", value = "indicator", 2:3) %>%
  group_by(PersonID) %>%
  summarise(pseo = sum(indicator)) %>%
  mutate(pseo.participant = ifelse(pseo > 0, 1, 0)) %>%
  ungroup() %>%
  select(PersonID, pseo.participant)

# Special Ed status

k12.enr.spced.indicator <- enrollment.original %>%
  select(PersonID, SpecialEdStatus) %>%
  mutate(SpecialEdStatus.code = ifelse(SpecialEdStatus %in% c(4,5,6,7,8,9), 1, 0)) %>%
  group_by(PersonID) %>%
  summarise(SpecialEdStatus.sum = sum(SpecialEdStatus.code)) %>%
  ungroup() %>%
  mutate(SpecialEdStatus = ifelse(SpecialEdStatus.sum > 0, 1, 0)) %>%
  select(PersonID, SpecialEdStatus)

# Home Primary Language

k12.enr.homeprimlang.indicator <- enrollment.original %>%
  select(PersonID, HomePrimaryLanguage) %>%
  mutate(HomePrimaryLanguage.code = ifelse(HomePrimaryLanguage == 11, 0, 1)) %>%
  group_by(PersonID) %>%
  summarise(HomePrimaryLanguage.sum = sum(HomePrimaryLanguage.code)) %>%
  ungroup() %>%
  mutate(non.english.home = ifelse(HomePrimaryLanguage.sum > 0, 1, 0)) %>%
  select(PersonID, non.english.home)

# Race

k12.enr.race <- enrollment.original %>%
  select(PersonID, RaceEthnicity) %>%
  mutate(RaceEthnicity = ifelse(RaceEthnicity == 0, "Unknown",
                                ifelse(RaceEthnicity == 1, "AI",
                                       ifelse(RaceEthnicity == 2, "Asian/PI",
                                              ifelse(RaceEthnicity == 3, "Hispanic",
                                                     ifelse(RaceEthnicity == 4, "Black", "White"))))),
         RaceEthnicity = as.factor(RaceEthnicity)) %>%
  distinct(PersonID, RaceEthnicity) %>%
  mutate(code = ifelse(RaceEthnicity == "Unknown", 0, 1)) %>%
  group_by(PersonID) %>%
  filter(code == max(code)) %>%
  ungroup %>%
  distinct(PersonID, .keep_all = TRUE) %>%
  select(PersonID, RaceEthnicity)

```

<br>

Now we can combine the list of graduates with the status indicators.

<br>

```{r k12enrollment combining all data}

k12.enrollment.master <- graduates %>%
  left_join(k12.enr.esi.indicator, by = "PersonID") %>%
  left_join(k12.enr.pseo.indicator, by = "PersonID") %>%
  left_join(k12.enr.spced.indicator, by = "PersonID") %>%
  left_join(k12.enr.homeprimlang.indicator, by = "PersonID") %>%
  left_join(k12.enr.race, by = "PersonID")

head(k12.enrollment.master)

names(k12.enrollment.master)

write_csv(k12.enrollment.master, "Data/SLEDS/K12 Enrollment/K12Enrollment-master.csv")
```

<br>

The prepped enrollment data has `r comma(nrow(k12.enrollment.master), accuracy = 1)` observations and `r comma(ncol(k12.enrollment.master), accuracy = 1)` columns.

<br>

# Summary of K12Enrollment data

Below is the percentage of total observations that;

* Were eligible for free or reduced lunch at some point in the dataset
* Are Female
* Have been homeless at some point in the dataset
* Had limited english proficiency at some point in the dataset
* Had non english speakders in home at some point in the dataset
* Participated in a PSEO course at some point in the dataset.
* Pct non white

As expected, nearly half (49.4%) of the observations are female. Over a third (39.2%) of the observations have been eligible for free or reduced lunch at some point in the dataset. Also, nearly a third (34.7%) of observations participated in a PSEO course. 8.7% of observations indicated that there was a non-english speaker at home, 2% had limited english proficiency and 0.4% indicated that were homeless at some point.

<br>

```{r prep k12enrollment summary, include=FALSE}
k12.enr.pct.total <- k12.enrollment.master %>%
  mutate(Gender = ifelse(Gender == "F", 1, 0),
         LimitedEnglishProficiencyIndicator = ifelse(LimitedEnglishProficiencyIndicator == "Y", 1, 0),
         HomelessIndicator = ifelse(HomelessIndicator == "Y", 1, 0),
         RaceEthnicity = ifelse(RaceEthnicity == "White", 0, 1)) %>%
  rename(Female = 4,
         `Limited English Proficiency` = 5,
         Homeless = 6,
         `Free or reduced lunch` = 7,
         `PSEO participant` = 8,
         `Special Ed` = 9,
         `Non-english in home` = 10,
         `Non-white` = 11) %>%
  gather(key = "indicator", value = "value", 4:11) %>%
  group_by(indicator) %>%
  summarise(confirmed.indicator = sum(value),
            total.observations = n()) %>%
  ungroup() %>%
  mutate(pct = scales::percent(confirmed.indicator / total.observations, accuracy = .1))
  
```

```{r table k12enrollment summary, echo=FALSE}
datatable(k12.enr.pct.total, class = "cell-border stripe", filter = "top", rownames = FALSE)

```

<br>

