---
title: "SAT"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(readxl)
library(janitor)
library(lubridate)
library(systemfonts)
reset_font_cache()
library(ggtext)
library(gmodels)
```

```{r themes and join docs, include=FALSE}
theme_line <- theme_bw() +
  theme(legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.text = element_text(margin = margin(l = 2)),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey70", size = 0.1),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))

theme_sf <- theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "white"),
        panel.border = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial"))

theme_bar <- theme_bw() +
  theme(panel.grid.major = element_line(color = "grey70", size = 0.1),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.key.size = unit(1, "lines"),
        legend.margin = margin(0,0,0,0),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        text = element_text(family = "Arial"))


mn_counties <- st_read("Data/Shapefiles/county shapefiles/MNCounties_MNDOT.shp", quiet = TRUE) %>%
  ms_simplify(keep = .01, keep_shapes = TRUE) %>%
  rename(countyfp = FIPS_CODE)

counties.regions <- read_csv("Data/Join docs/county_regions.csv") %>%
  rename(mif = `MIF Region`) %>%
  mutate(countyfp = formatC(countyfp, width = 3, flag = "0"),
         Name = str_to_title(Name),
         Name = str_replace(Name, "Q", "q"),
         Name = str_replace(Name, "Of The", "of the"),
         Name = str_replace(Name, "Mcleod", "McLeod"),
         Dem_Desc = ifelse(Name == "Minnesota", "Minnesota", Dem_Desc) ,
         edr = str_replace(edr, "  ", " "),
         planning.region = str_replace(planning.region, " Minnesota", ""),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"),
         mif = ifelse(is.na(mif), "TC", mif),
         mif = as.factor(mif),
         mif = fct_relevel(mif, "NW", "NE", "WC", "EC", "SW", "SE", "TC"))

```

<br>

# Data prep - SAT

Alright, lets see what this dataset looks like.

<br>

```{r sat original}
sat.original <- read_csv("Data/SLEDS/SAT/sat.csv") %>%
  mutate(PersonID = as.integer(PersonID))

head(sat.original)

names(sat.original)
```

<br>

There are `r comma(sat.original %>% nrow(), accuracy = 1)` rows and `r comma(sat.original %>% ncol(), accuracy = 1)` columns. The definitions of the columns are;

* SATTimeID: Time of the test
* TestType: SAT
* ExamSubject: subject
* ExamGrade: Exam grades are all over the place here and there's no definition in the data dictionary.

Okay, since I don't have any information related to how the grading system works, I will just utilize this dataset to code wither a PersonID has taken an SAT test.

<br>

```{r sat taken}
sat.taken <- sat.original %>%
  mutate(sat.taken = 1) %>%
  select(PersonID, sat.taken) %>%
  distinct(PersonID, .keep_all = TRUE)

head(sat.taken)

names(sat.taken)
```

<br>

Now we will join it with the master dataset.

<br>

```{r sat join with master}
master.8 <- read_csv("Data/SLEDS/Masters/Master-7.csv") %>%
  mutate(PersonID) %>%
  left_join(sat.taken, by = "PersonID") %>%
  mutate(sat.taken = ifelse(is.na(sat.taken), 0, sat.taken))

head(master.8)

names(master.8)
```

<br>

As expected there are `r comma(master.8 %>% nrow(), accuracy = 1)` rows and `r comma(master.8 %>% ncol(), accuracy = 1)` columns.


# Summary of SAT data

Lets check to see what percentage of the master dataset has taken the SAT.

There is a tiny percentage of the dataset that has taken the SAT, 0.6%. With this information I don't see a need to check whether there is a significant difference in students taking the SAT based on RUCA categoriy or EDR. Even if there is a significant difference, there just isn't enough observations to make a difference.

<br>

```{r pct SAT total}
sat.total <- master.8 %>%
  group_by(sat.taken) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(pct = n / sum(n))

datatable(sat.total, class = "cell-border stripe", filter = "top", rownames = FALSE,
          options = list(columnDefs = list(list(className = "dt-center", targets = 0:1))))  %>%
  formatPercentage(3, digits = 1) %>%
  formatCurrency(2, digits = 0, "")
```

```{r write master}
write_csv(master.8, "Data/SLEDS/Masters/Master-8.csv")
```

