---
title: "Visualizations - local employment cross tabs"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
library(htmlwidgets)
library(RColorBrewer)
library(readxl)
library(janitor)
library(lubridate)
library(systemfonts)
reset_font_cache()
library(ggtext)
library(gmodels)
library(fastDummies)
library(car)
library(glmnet)
library(glmnetUtils)
library(pscl)
library(sjPlot)
library(ggforce)

```

```{r join docs, include=FALSE}
theme_bar <- theme_bw() +
  theme(panel.grid.major = element_line(color = "grey70", size = 0.1),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.key.size = unit(1, "lines"),
        legend.margin = margin(0,0,0,0),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        text = element_text(family = "Arial") ,
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"))

theme_line <- theme_bw() +
  theme(legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.text = element_text(margin = margin(l = 2)),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey70", size = 0.1),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial") ,
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"))


theme_sf <- theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "white"),
        panel.border = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(margin = margin(l = 2)),
        legend.margin = margin(0,0,0,0),
        legend.key.size = unit(1, "lines"),
        text = element_text(family = "Arial") ,
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"))

regions <- read_csv("Data/Join docs/county_regions.csv") %>%
    select(5,6) %>%
    unique() %>%
    mutate(edr = str_replace(edr, "  ", " "),
           planning.region = str_replace(planning.region, " Minnesota", ""),
           planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
           edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"))

counties.regions <- read_csv("Data/Join docs/county_regions.csv") %>%
  rename(mif = `MIF Region`) %>%
  mutate(countyfp = formatC(countyfp, width = 3, flag = "0"),
         Name = str_to_title(Name),
         Name = str_replace(Name, "Q", "q"),
         Name = str_replace(Name, "Of The", "of the"),
         Name = str_replace(Name, "Mcleod", "McLeod"),
         Dem_Desc = ifelse(Name == "Minnesota", "Minnesota", Dem_Desc) ,
         edr = str_replace(edr, "  ", " "),
         planning.region = str_replace(planning.region, " Minnesota", ""),
         planning.region = fct_relevel(planning.region, "Northwest", "Northeast", "Central", "Seven County Mpls-St Paul", "Southwest", "Southeast"),
         edr = fct_relevel(edr, "EDR 1 - Northwest", "EDR 2 - Headwaters", "EDR 3 - Arrowhead", "EDR 4 - West Central", "EDR 5 - North Central", "EDR 6E- Southwest Central", "EDR 6W- Upper Minnesota Valley", "EDR 7E- East Central", "EDR 7W- Central", "EDR 8 - Southwest", "EDR 9 - South Central", "EDR 10 - Southeast", "EDR 11 - 7 County Twin Cities", "Minnesota"),
         mif = ifelse(is.na(mif), "TC", mif),
         mif = as.factor(mif),
         mif = fct_relevel(mif, "NW", "NE", "WC", "EC", "SW", "SE", "TC"))


color.ruca <- c("Entirely rural" = "#009933", "Town/rural mix" = "#99CC33", "Urban/town/rural mix" = "#CC9966", "Entirely urban" = "#754C29", "Minnesota" = "black")

color.pr <- c("Northwest" = 	"#4575b4", "Northeast" = "grey", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black")

color.edr <- c("EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365", "Minnesota" = "black")

color.pr.edr <- c ("Northwest" = "#4575b4","Northeast" = "#e0f3f8", "Central" = "#fee090", "Seven County Mpls-St Paul" = "#d73027", "Southwest" = "#91bfdb", "Southeast" = "#fc8d59", "Minnesota" = "black", "EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365")

mn_counties <- st_read("Data/Shapefiles/county shapefiles/MNCounties_MNDOT.shp", quiet = TRUE) %>%
  ms_simplify(keep = .01, keep_shapes = TRUE) %>%
  rename(countyfp = FIPS_CODE)
```

```{r master}
master <- read_csv("Data/SLEDS/Masters/Master-after-ct.csv")

dep.var.names <- c("hs.grad.year.county.match", "hs.grad.year.2.county.match", "hs.grad.year.7.county.match", "hs.grad.year.edr.match", "hs.grad.year.2.edr.match", "hs.grad.year.7.edr.match", "hs.grad.year.region.match", "hs.grad.year.region.match", "hs.grad.year.2.region.match", "hs.grad.year.7.region.match", "hs.grad.year.state.match", "hs.grad.year.2.state.match", "hs.grad.year.7.state.match")

```

<br>

From the cross tabs analysis, we've concluded that the following variables are important;

```{r ind var}
master %>%
  select(-1:-13) %>%
  names() 
```

However, there are plenty of variables here that are VERY similar. For example, whether an individual graduated from a post-secondary institution and the type of institution they graduated from. So let's simplify the variables to what is most important and has the least amount of overlap with other variables.

* Demographic
    + RaceEthnicity
* High school characteristics
    + Dem_Desc
    + EDR
    + avg.wages.pct.state
* High school enrollment
    + LimitedEnglishProficiencyIndicator
    + PSEO.participant
* High school accomplishments
    + took.act
    + ap.exam
    + cte.achievement
    + MCA.M
    + MCA.R
    + MCA.S
* Post-secondary
    + ps.grad.InstitutionSector
    + highest.cred.level
    
We eliminated the following variables for now;

* non.english.home
* english.learner
* total.cte.courses.taken
* ps.grad
* attended.ps

In addition to eliminating those variables, we will also simplify the ps.grad.InstitutionSector to the following codes;

* 1 = graduated from the following insitutions
    + Public, 4-year (1)
    + Private, not for profit 4-year (2)
    + Private, for profit 4-year (3)
* 2 = graduated from the following institutions
    + Private, not for profit 2-year (5, 11)
    + Private, for profit 2-year (6, 11)
    + Public, less than 2-year (7, 11)
    + Private, not for profit less than 2-year (8, 11)
    + Private for-profit, less-than 2-year (9, 11)
* 3 = graduated from multiple institution sectors (10)
* 4 = graduated from a public, 2-year institution (4,4)
* 5 = Did not graduate or attend a post-secondary institution

```{r updating master}
updated.master <- master %>%
  select(`dep.var.names`, RaceEthnicity, Dem_Desc, edr, avg.wages.pct.state, LimitedEnglishProficiencyIndicator, pseo.participant, took.ACT, ap.exam, cte.achievement, MCA.M, MCA.R, MCA.S, ps.grad.InstitutionSector, highest.cred.level) %>%
  mutate(ps.grad.InstitutionSector = ifelse(ps.grad.InstitutionSector %in% c(1,2,3), 1,
                                            ifelse(ps.grad.InstitutionSector == 11, 2,
                                                   ifelse(ps.grad.InstitutionSector == 10, 3,
                                                          ifelse(ps.grad.InstitutionSector == 4, 4, 5))))) 

updated.ind.var.names <- updated.master %>%
  select(-`dep.var.names`) %>%
  names()

```

<br>

## Demographic variables{.tabset}

**RaceEthnicity:** The chart shows that for grad year +0 and grad year +2, a significantly larger proportion of Black and Hispanic individuals have meaningful employment across all geographies. However, by grad year +7 that difference between race and ethnicity doesn't exist and the proportions are more equal.

<br>

### RaceEthnicity

```{r demographic race ethnicity vis}
dem.race.ethnicity.vis <- updated.master %>%
  select(`dep.var.names`, RaceEthnicity) %>%
  gather(key = "grad.year", value = "meaningful.emp", 1:12) %>%
  filter(meaningful.emp != "Unknown") %>%
  filter(meaningful.emp != "After 2019") %>%
  filter(RaceEthnicity != "Unknown") %>%
  mutate(geography = str_sub(grad.year, -12, -7),
         geography = str_replace(geography, ".2.edr", "EDR"),
         geography = str_replace(geography, ".7.edr", "EDR"),
         geography = str_replace(geography, ".state", "MN"),
         geography = str_replace(geography, "ar.edr", "EDR"),
         geography = str_replace(geography, "region", "All 3 EDRs"),
         geography = str_replace(geography, "county", "County"),
         geography = as.factor(geography),
         geography = fct_relevel(geography,"County", "EDR", "All 3 EDRs", "MN"),
         grad.year = str_sub(grad.year, 14, 14),
         grad.year = str_replace(grad.year, "c", "0"),
         grad.year = str_replace(grad.year, "e", "0"),
         grad.year = str_replace(grad.year, "r", "0"),
         grad.year = str_replace(grad.year, "s", "0"),
         grad.year = paste("Grad year +", grad.year, sep = ""),
         grad.year = as.factor(grad.year)) %>%
  group_by(grad.year, geography, meaningful.emp, RaceEthnicity) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(grad.year, geography, RaceEthnicity) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  filter(meaningful.emp == "Location match") %>%
  mutate(data_id = seq(n()))

dem.race.ethnicity.plot <- ggplot(dem.race.ethnicity.vis, aes(geography, pct, fill = RaceEthnicity, group = RaceEthnicity)) +
  facet_wrap(~grad.year, ncol = 2, scales = "free_y") +
  geom_col_interactive(position = "dodge", aes(data_id = data_id, tooltip = paste(RaceEthnicity, "\nTime X: ", grad.year, "\nNumber of individuals with meaningful employment: ", comma(n), "\nProportion of race and ethnicity: ", percent(pct, accuracy = .1), sep = ""))) +
  geom_mark_ellipse(data = filter(dem.race.ethnicity.vis, RaceEthnicity == "Hispanic" & grad.year %in% c("Grad year +0", "Grad year +2"))) +
  geom_mark_ellipse(data = filter(dem.race.ethnicity.vis, RaceEthnicity == "Black" & grad.year %in% c("Grad year +0", "Grad year +2"))) +
  labs(x="", y = "", color="", title = "Proportion of race and ethnicity with meaningful employment at each\nTime X ") + 
  geom_text(data = filter(dem.race.ethnicity.vis, RaceEthnicity == "Hispanic" & grad.year %in% c("Grad year +0")), aes(x = 1.5, y = .27, label = "A larger proportion of\nHispanic and Black individuals\nhave meaningful employment"), size = 5) +
  geom_text(data = filter(dem.race.ethnicity.vis, RaceEthnicity == "Hispanic" & grad.year %in% c("Grad year +2")), aes(x = 1.5, y = .4, label = "A larger proportion of\nHispanic and Black\nindividuals have meaningful\nemployment"), size = 5) +
   geom_text(data = filter(dem.race.ethnicity.vis, RaceEthnicity == "Hispanic" & grad.year %in% c("Grad year +7")), aes(x = 2, y = .4, label = "The proportions are more\nequal across Race\nand Ethnicity"), size = 5) +
  scale_y_continuous(labels=scales::percent)+
  theme_bar+
  scale_fill_manual(values = brewer.pal(n = 5, "RdYlBu"),
                    guide = guide_legend(ncol = 3)) +
  theme(legend.position = "bottom",
        text = element_text(size = 18))


girafe(ggobj = dem.race.ethnicity.plot, width_svg = 12, height_svg = 15) %>%
  girafe_options(opts_selection(type = "none"),
                 opts_sizing(rescale = FALSE))      

```

<br>

## High school characteristics variables{.tabset}

Summary

<br>

### Dem_Desc

```{r high school characterstics Dem_Desc vis}
hscharact.dem.desc.vis <- updated.master %>%
  select(`dep.var.names`, Dem_Desc) %>%
  gather(key = "grad.year", value = "meaningful.emp", 1:12) %>%
  filter(meaningful.emp != "Unknown") %>%
  filter(meaningful.emp != "After 2019") %>%
  filter(Dem_Desc != "Unknown") %>%
  mutate(geography = str_sub(grad.year, -12, -7),
         geography = str_replace(geography, ".2.edr", "EDR"),
         geography = str_replace(geography, ".7.edr", "EDR"),
         geography = str_replace(geography, ".state", "MN"),
         geography = str_replace(geography, "ar.edr", "EDR"),
         geography = str_replace(geography, "region", "All 3 EDRs"),
         geography = str_replace(geography, "county", "County"),
         geography = as.factor(geography),
         geography = fct_relevel(geography,"County", "EDR", "All 3 EDRs", "MN"),
         grad.year = str_sub(grad.year, 14, 14),
         grad.year = str_replace(grad.year, "c", "0"),
         grad.year = str_replace(grad.year, "e", "0"),
         grad.year = str_replace(grad.year, "r", "0"),
         grad.year = str_replace(grad.year, "s", "0"),
         grad.year = paste("Grad year +", grad.year, sep = ""),
         grad.year = as.factor(grad.year),
         Dem_Desc = fct_relevel(Dem_Desc, "Entirely rural", "Town/rural mix", "Urban/town/rural mix")) %>%
  group_by(grad.year, geography, meaningful.emp, Dem_Desc) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(grad.year, geography, Dem_Desc) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  filter(meaningful.emp == "Location match") %>%
  mutate(data_id = seq(n()))

hscharact.dem.desc.plot <- ggplot(hscharact.dem.desc.vis, aes(geography, pct, fill = Dem_Desc, group = Dem_Desc)) +
  facet_wrap(~grad.year, ncol = 2, scales = "free_y") +
  geom_col_interactive(position = "dodge", aes(data_id = data_id, tooltip = paste(Dem_Desc, "\nTime X: ", grad.year, "\nNumber of individuals with meaningful employment: ", comma(n), "\nProportion of RUCA category: ", percent(pct, accuracy = .1), sep = ""))) +
  geom_mark_ellipse(data = filter(hscharact.dem.desc.vis, Dem_Desc == "Town/rural mix" & grad.year %in% c("Grad year +0", "Grad year +2"))) +
  geom_label(data = filter(hscharact.dem.desc.vis, grad.year %in% c("Grad year +0")), aes(x = 1.5, y = .2, label = "A larger proportion of\nindividuals from town/rural mix\nhigh schools have meaningful\nemployment"), size = 5, fill = "#fdae61") +
  geom_label(data = filter(hscharact.dem.desc.vis, grad.year %in% c("Grad year +2")), aes(x = 1.5, y = .32, label = "A larger proportion of\nindividuals from town/rural mix\nhigh schools have meaningful\nemployment"), size = 5, fill = "#fdae61") +
  geom_mark_ellipse(data = filter(hscharact.dem.desc.vis, Dem_Desc == "Urban/town/rural mix" & grad.year %in% c("Grad year +7") & geography %in% c("County", "EDR", "All 3 EDRs"))) +
   geom_label(data = filter(hscharact.dem.desc.vis, grad.year %in% c("Grad year +7")), aes(x = 2, y = .4, label = "The proportions of individuals\nfrom urban/town/rural mix high\nschools are significantly lower"), size = 5, fill = "#ffffbf") +
  labs(x="", y = "", color="", title = "Proportion of individuals graduating from RUCA categories with meaningful\nemployment at each Time X ") + 
  scale_y_continuous(labels=scales::percent)+
  theme_bar+
  scale_fill_manual(values = brewer.pal(n = 5, "RdYlBu"),
                    guide = guide_legend(ncol = 3)) +
  theme(legend.position = "bottom",
        text = element_text(size = 18))


girafe(ggobj = hscharact.dem.desc.plot, width_svg = 12, height_svg = 15) %>%
  girafe_options(opts_selection(type = "none"),
                 opts_sizing(rescale = FALSE))      

```